\defclass{ChronoSingleThread}

\textbf{This class is deprecated} but kept for
compatibility with older versions of SSJ.
\class{Chrono} should be used instead of \class{ChronoSingleThread}.
The \class{ChronoSingleThread} class extends the \class{AbstractChrono} 
class and computes the CPU time for the current thread only.
This is the simplest way to use chronos. Classes \class{AbstractChrono},
\class{SystemTimeChrono}, \class{GlobalCPUTimeChrono} and 
\class{ThreadCPUTimeChrono} provide different chronos implementations
(see these classes to learn more about SSJ chronos).


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bigskip\hrule

\begin{code}
\begin{hide}
/*
 * Class:        ChronoSingleThread
 * Description:  deprecated
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.util;\begin{hide}

import java.lang.management.ManagementFactory;
import java.lang.management.ThreadMXBean;
\end{hide}


@Deprecated
public class ChronoSingleThread \begin{hide} extends AbstractChrono {

   private ThreadCPUTimeChrono chrono = new ThreadCPUTimeChrono();

   protected void getTime (long[] tab) {
         chrono.getTime(tab);
   }\end{hide}
\end{code}

%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Constructor}

\begin{code}

   public ChronoSingleThread()\begin{hide} {
      chrono.init();
      init();
   }\end{hide}
\end{code}
  \begin{tabb} Constructs a \texttt{ChronoSingleThread} object and
    initializes it to zero.
  \end{tabb}
\begin{code}
\begin{hide}
}\end{hide}
\end{code}
