\defclass{ClassFinder}

Utility class used to convert a simple class name to
a fully qualified class object.
The \externalclass{java.lang}{Class} class can be used to obtain
information about a class (its name, its fields, methods,
constructors, etc.), and to construct objects, even
if the exact class is known at runtime only.
It provides a \externalmethod{java.lang}{Class}{forName}{(String)} static method converting
a string to a \externalclass{java.lang}{Class}, but the given string
must be a fully qualified name.

Sometimes, configuration files may need
to contain Java class names. After they are
extracted from the file, these class names are
given to \externalmethod{java.lang}{Class}{forName}{(String)} to be converted into
\externalclass{java.lang}{Class} objects.
Unfortunately, only
fully qualified class names will be accepted
as input, which clutters
configuration files, especially if long package names are used.
This class permits the definition of a set of
import declarations in a way similar to the
Java Language Specification \cite{iGOS00a}.
It provides methods to convert a simple class name
to a \externalclass{java.lang}{Class} object and to generate a simple name
from a \externalclass{java.lang}{Class} object, based on the import
rules.

The first step for using a class finder is to construct an instance of
this class.
Then, one needs to retrieve the initially empty list of import
declarations by
using \method{getImports}{}, and update it with the
actual import declarations.
Then, the method \method{findClass}{} can find a class using the
import declarations.
For example, the following code retrieves the class object for the
\texttt{List} class in package \texttt{java.util}
\begin{vcode}
   ClassFinder cf = new ClassFinder();
   cf.getImports().add ("java.util.*");
   Class<?> listClass = cf.findClass ("List");
\end{vcode}

\bigskip\hrule\bigskip

\begin{code}\begin{hide}
/*
 * Class:        ClassFinder
 * Description:  Convert a simple class name to a fully qualified class object
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.util;\begin{hide}

import java.util.ArrayList;
import java.util.LinkedList;
import java.util.List;
\end{hide}


public class ClassFinder implements Cloneable, java.io.Serializable\begin{hide} {
   private static final long serialVersionUID = -4847630831331065792L;
\end{hide}

   private List<List<String>> imports\begin{hide} = new LinkedList<List<String>>();
\end{hide}
\end{code}
\begin{tabb}   Contains the saved import lists.
 Each element of this list is a nested \externalclass{java.util}{List} containing
 \externalclass{java.lang}{String}'s, each string containing the fully qualified
 name of an imported package or class.
\end{tabb}
\begin{htmlonly}
   \blocktag{@serial}{}
\end{htmlonly}
\begin{code}

   public ClassFinder()\begin{hide} {
      List<String> imp = new ArrayList<String>();
      imports.add (imp);
   }\end{hide}
\end{code}
\begin{tabb}   Constructs a new class finder with
 an empty list of import declarations.
\end{tabb}
\begin{code}

   public List<String> getImports()\begin{hide} {
      return imports.get (imports.size() - 1);
   }\end{hide}
\end{code}
\begin{tabb}   Returns the current list of import declarations.
 This list may contain only \externalclass{java.lang}{String}'s
 of the form \texttt{java.class.name} or
 \texttt{java.package.name.*}.
\end{tabb}
\begin{htmlonly}
   \return{the current list of import declarations.}
\end{htmlonly}
\begin{code}

   public void saveImports()\begin{hide} {
      List<String> imp = getImports();
      List<String> impBack = new ArrayList<String> (imp);
      imports.add (impBack);
   }\end{hide}
\end{code}
\begin{tabb}   Saves the current import list on the import stack.
 This method makes a copy of the list returned
 by \method{getImports}{()} and puts it on top
 of a stack to be restored later by
 \method{restoreImports}{()}.
\end{tabb}
\begin{code}

   public void restoreImports()\begin{hide} {
      if (imports.size() == 1)
         getImports().clear();
      else
         imports.remove (imports.size() - 1);
   }\end{hide}
\end{code}
\begin{tabb}   Restores the list of import declarations.
 This method removes the last list of import
 declarations from the stack. If the stack
 contains only one list, this list is cleared.
\end{tabb}
\begin{code}

   public Class<?> findClass (String name) throws
      ClassNotFoundException, NameConflictException\begin{hide} {
      // Try to consider the name as a fully qualified class name
      try {
         return Class.forName (name);
      }
      catch (ClassNotFoundException cnfe) {}

      List<String> imports = getImports();
      Class<?> candidate = null;
      String candidateImportString = "";
      boolean candidateImportOnDemand = false;

      // Determines the name of the outermost class
      // if name corresponds to the simple name of a nested class, e.g., for A.B.C
      // outerName will contain A.
      int idxOut = name.indexOf ('.');
      String outerName;
      if (idxOut == -1)
         outerName = name;
      else
         outerName = name.substring (0, idxOut);
      for (String importString : imports) {
         // For each import declaration, we try to load
         // a class and store the result in cl.
         // When cl is not null, the Class object is
         // compared with the best candidate candidate.
         Class<?> cl = null;
         boolean onDemand = false;
         if (!importString.endsWith (".*")) {
            // Single-type import declaration
            // We must ensure that this will correspond to
            // the desired class name. For example, if we
            // search List and the import java.util.ArrayList
            // is found in importString, the ArrayList
            // class must not be returned.
            if (importString.endsWith ("." + outerName)) {
               // The name of outer class was found in importString and
               // a period is found left to it.
               // So try to load this class.
               // Simple class names have precedence over
               // on-demand names.

               // Replace, in importString, the name of
               // the outer class with the true class name
               // we want to load.  If the simple name
               // does not refer to a nested class, this
               // has no effect.
               String cn = importString.substring
                  (0, importString.length() - outerName.length()) + name;
               try {
                  cl = Class.forName (cn);
               }
               catch (ClassNotFoundException cnfe) {}
            }
         }
         else {
            // Type import on demand declaration
            try {
               // Replace the * with name and
               // try to load the class.
               // If that succeeds, our candidate cl
               // is onDemand.
               cl = Class.forName
                  (importString.substring (0, importString.length() - 1) + name);
               onDemand = true;
            }
            catch (ClassNotFoundException cnfe) {}
         }
         if (cl != null) {
            // Something was loaded
            if (candidate == null ||
                (candidateImportOnDemand && !onDemand)) {
               // We had no candidate or the candidate was imported
               // on-demand while this one is a single-type import.
               candidate = cl;
               candidateImportString = importString;
               candidateImportOnDemand = onDemand;
            }
            else if (candidate != cl)
               throw new NameConflictException
                  (this, name,
                   "simple class name " + name +
                   " matches " + candidate.getName() +
                   " (import string " + candidateImportString + ") or " +
                   cl.getName() + " (import string " + importString + ")");
         }
      }
      if (candidate == null)
         throw new ClassNotFoundException
            ("Cannot find the class with name " + name);
      return candidate;
   }\end{hide}
\end{code}
\begin{tabb}   Tries to find the class corresponding to the simple
 name \texttt{name}.  The method first considers
 the argument as a fully qualified class name
 and calls \externalmethod{java.lang}{Class}{forName}{(String)} \texttt{(name)}.
 If the class cannot be found, it considers the argument
 as a simple name.  A simple name refers to a class without
 specifying the package declaring it.  To convert
 simple names to qualified names, the method iterates through
 all the strings in the list returned by \method{getImports}{()},
 applying
 the same rules as a Java compiler to resolve
 the class name.  However, if an imported package or
 class does not exist, it will be ignored whereas
 the compiler would stop with an error.

 For the class with simple name \texttt{name} to
 be loaded, it must be imported explicitly (single-type import) or
 one of the imported packages must contain it (type import on-demand).
 If the class with name \texttt{name} is imported explicitly,
 this import declaration has precedence over
 any imported packages.
 If several import declaration match the given simple
 name, e.g., if several fully qualified names with the same
 simple name are imported, or if a class with simple
 name \texttt{name} exists in several packages,
 a \class{NameConflictException} is thrown.
\end{tabb}
\begin{htmlonly}
   \param{name}{the simple name of the class.}
   \return{a reference to the class being loaded.}
   \exception{ClassNotFoundException}{if the class
    cannot be loaded.}
   \exception{NameConflictException}{if a name conflict occurred.}
\end{htmlonly}
\begin{code}

   public String getSimpleName (Class<?> cls)\begin{hide} {
      if (cls.isArray())
         return getSimpleName (cls.getComponentType()) + "[]";
      if (cls.isPrimitive())
         return cls.getName();
      Class<?> outer = cls;
      while (outer.getDeclaringClass() != null)
         outer = outer.getDeclaringClass();
      boolean needsFullyQualified = true;
      for (String importString : getImports()) {
         if (importString.equals (outer.getName()))
             // A single-type import is given, can return an unqualified name.
            needsFullyQualified = false;
         else if (importString.endsWith (".*")) {
            // Remove the .* at the end of the import string to get a package name.
            String pack = importString.substring (0, importString.length() - 2);
            // Compare this package name.
            if (pack.equals (cls.getPackage().getName()))
               needsFullyQualified = false;
         }
      }
      if (needsFullyQualified)
         return cls.getName();
      else {
         String name = cls.getName();
         String pack = cls.getPackage().getName();
         if (!name.startsWith (pack))
            throw new IllegalStateException
               ("The class name " + name +
                " does not contain the package name " + pack);

         // Removes the package and the . from the fully qualified class name.
         return name.substring (pack.length() + 1);
      }         
   }\end{hide}
\end{code}
\begin{tabb}   Returns the simple name of the class \texttt{cls} that
 can be used when the imports contained
 in this class finder are used.
 For example, if \texttt{java.lang.String.class} is given
 to this method, \texttt{String} is returned if
 \texttt{java.lang.*} is among the import declarations.

 Note: this method does not try to find name conflicts.
 This operation is performed by \method{findClass}{(String)} only.
 For example, if the list of imported declarations
 contains \texttt{foo.bar.*} and \texttt{test.Foo}, and
 the simple name for \texttt{test.Foo} is queried,
 the method returns \texttt{Foo} even if the package
 \texttt{foo.bar} contains a \texttt{Foo} class.
\end{tabb}
\begin{htmlonly}
   \param{cls}{the class for which the simple name is queried.}
   \return{the simple class name.}
\end{htmlonly}
\begin{code}

   public ClassFinder clone()\begin{hide} {
      ClassFinder cf;
      try {
         cf = (ClassFinder)super.clone();
      }
      catch (CloneNotSupportedException cne) {
         throw new InternalError
            ("CloneNotSupported thrown for a class implementing Cloneable");
      }
      cf.imports = new LinkedList<List<String>>();
      for (List<String> imp : imports) {
         List<String> impCpy = new ArrayList<String> (imp);
         cf.imports.add (impCpy);
      }
      return cf;
   }
}\end{hide}
\end{code}
\begin{tabb}   Clones this class finder, and copies its lists of
  import declarations.
\end{tabb}
