\defclass{NameConflictException}

This exception is thrown by a \class{ClassFinder}
when two or more fully qualified class names can be
associated with a simple class name.

\bigskip\hrule
\begin{code}\begin{hide}
/*
 * Class:        NameConflictException
 * Description:  An exception 
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */\end{hide}
package umontreal.iro.lecuyer.util;


public class NameConflictException extends Exception\begin{hide} {
   private static final long serialVersionUID = -5124156035520217708L;
   private ClassFinder finder;
   private String name;
\end{hide}
\end{code}

%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection* {Constructors}
\begin{code} 

   public NameConflictException()\begin{hide} {}\end{hide}
\end{code}
\begin{tabb}   Constructs a new name conflict exception.
\end{tabb}
\begin{code}

   public NameConflictException (String message)\begin{hide} {
      super (message);
   }\end{hide}
\end{code}
\begin{tabb}   Constructs a new name conflict exception with
 message \texttt{message}.
\end{tabb}
\begin{htmlonly}
   \param{message}{the error message.}
\end{htmlonly}
\begin{code}

   public NameConflictException (ClassFinder finder, String name,
                                 String message)\begin{hide} {
      super (message);
      this.finder = finder;
      this.name = name;
   }\end{hide}
\end{code}
\begin{tabb}   Constructs a new name conflict exception with class
 finder \texttt{finder}, simple name \texttt{name},
 and message \texttt{message}.
\end{tabb}
\begin{htmlonly}
   \param{finder}{the class finder in which the name conflict occurred.}
   \param{name}{the simple conflicting name.}
   \param{message}{the message describint the conflict.}
\end{htmlonly}

%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection* {Methods}
\begin{code}

   public ClassFinder getClassFinder()\begin{hide} {
      return finder;
   }\end{hide}
\end{code}
\begin{tabb}   Returns the class finder associated with
 this exception.
\end{tabb}
\begin{htmlonly}
   \return{the associated class finder.}
\end{htmlonly}
\begin{code}

   public String getName()\begin{hide} {
      return name;
   }
}\end{hide}
\end{code}
\begin{tabb}   Returns the simple name associated with
 this exception.
\end{tabb}
\begin{htmlonly}
   \return{the associated simple name.}
\end{htmlonly}
