\defclass{RatioFunction}

Represents a function computing a ratio of two values.

\bigskip\hrule

\begin{code}
\begin{hide}
/*
 * Class:        RatioFunction
 * Description:  Represents a function computing a ratio of two values.
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.util;\begin{hide}
\end{hide}


public class RatioFunction implements MultivariateFunction\begin{hide} {
   private double zeroOverZero = Double.NaN;
\end{hide}
\end{code}
\subsubsection*{Constructors}
\begin{code}

   public RatioFunction()\begin{hide} {}\end{hide}
\end{code}
\begin{tabb}   Constructs a new ratio function.
\end{tabb}
\begin{code}

   public RatioFunction (double zeroOverZero)\begin{hide} {
      this.zeroOverZero = zeroOverZero;
   }\end{hide}
\end{code}
\begin{tabb}   Constructs a new ratio function that returns
   \texttt{zeroOverZero} for the special case of $0/0$.
   See the \method{getZeroOverZeroValue}{}  method for more information.
   The default value of \texttt{zeroOverZero} is \texttt{Double.NaN}.
\end{tabb}
\begin{htmlonly}
   \param{zeroOverZero}{the value for $0/0$.}
\end{htmlonly}
\subsubsection*{Methods}
\begin{code}

   public double getZeroOverZeroValue()\begin{hide} {
      return zeroOverZero;
   }\end{hide}
\end{code}
\begin{tabb}   Returns the value returned by \method{evaluate}{(double[])} in the
 case where the $0/0$ function is calculated.
 The default value for $0/0$ is \texttt{Double.NaN}.

 Generally, $0/0$ is undefined, and therefore associated with the \texttt{Double.NaN}
  constant, meaning \emph{not-a-number}.
 However, in certain applications, it can be defined differently to accomodate some special cases.
 For exemple, in a queueing system, if there are no arrivals, no customers are served,
 lost, queued, etc.  As a result, many performance measures of interest turn out to be
 $0/0$.  Specifically, the loss probability, i.e., the ratio of
 lost customers over the number of arrivals, should be 0 if there is no arrival;
 in this case, $0/0$ means 0.
 On the other hand, the service level, i.e., the fraction of customers waiting less than
 a fixed threshold, could be fixed to 1 if there is no arrival.
\end{tabb}
\begin{htmlonly}
   \return{the value for $0/0$.}
\end{htmlonly}
\begin{code}

   public void setZeroOverZeroValue (double zeroOverZero)\begin{hide} {
      this.zeroOverZero = zeroOverZero;
   }\end{hide}
\end{code}
\begin{tabb}   Sets the value returned by \method{evaluate}{(double[])} for
 the undefined function $0/0$ to \texttt{zeroOverZero}.
 See \method{getZeroOverZeroValue}{} for more information.
\end{tabb}
\begin{htmlonly}
   \param{zeroOverZero}{the new value for $0/0$.}
\end{htmlonly}
\begin{code}\begin{hide}

   public int getDimension() {
      return 2;
   }

   public double evaluate (double... x) {
      if (x.length != 2)
         throw new IllegalArgumentException
            ("Invalid length of x");
      if (x[0] == 0 && x[1] == 0)
         return zeroOverZero;
      return x[0]/x[1];
   }

   public double evaluateGradient (int i, double... x) {
      if (x.length != 2)
         throw new IllegalArgumentException
            ("Invalid length of x");
      switch (i) {
      case 0: return 1.0/x[1];
      case 1: return -x[0]/(x[1]*x[1]);
      default: throw new IndexOutOfBoundsException
         ("Invalid value of i: " + i); 
      }
   }
}\end{hide}
\end{code}
