\defclass{TransformingList}

Represents a list that dynamically transforms the elements 
of another list.
This abstract class defines a list
containing an inner list of elements of a certain type,
and provides facilities to convert these
inner elements to outer elements of another
type.
A concrete subclass simply needs to provide
methods for converting between the inner
and the outer types.\begin{htmlonly}
   \param{<IE>}{the inner type of the elements.}
   \param{<OE>}{the type of the outer elements.}
\end{htmlonly}


\bigskip\hrule

\begin{code}
\begin{hide}
/*
 * Class:        TransformingList
 * Description:  List that dynamically transforms the elements of another list.
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.util;\begin{hide}

import java.util.AbstractList;
import java.util.List;
import java.util.Iterator;
import java.util.ListIterator;
\end{hide}


public abstract class TransformingList<OE,IE> extends AbstractList<OE>\begin{hide} {
   private List<IE> fromList;
\end{hide}

   public TransformingList (List<IE> fromList)\begin{hide} {
      this.fromList = fromList;
   }\end{hide}
\end{code}
\begin{tabb}   Creates a new transforming list wrapping
 the inner list \texttt{fromList}.
\end{tabb}
\begin{htmlonly}
   \param{fromList}{the inner list.}
\end{htmlonly}
\begin{code}\begin{hide}

   public List<IE> getInnerList() {
      return fromList;
   }\end{hide}

   public abstract OE convertFromInnerType (IE e)\begin{hide};\end{hide}
\end{code}
\begin{tabb} Converts an element in the inner list to
 an element of the outer type.
\end{tabb}
\begin{htmlonly}
   \param{e}{the inner element.}
   \return{the outer element.}
\end{htmlonly}
\begin{code}

   public abstract IE convertToInnerType (OE e)\begin{hide};\end{hide}
\end{code}
\begin{tabb} Converts an element of the outer type to
 an element for the inner list.
\end{tabb}
\begin{htmlonly}
   \param{e}{the outer element.}
   \return{the inner element.}
\end{htmlonly}
\begin{code}\begin{hide}

   public void add (int index, OE element) {
      IE fe = convertToInnerType (element);
      fromList.add (fe);
   }

   @Override
   public void clear () {
      fromList.clear();
   }

   @Override
   public OE get (int index) {
      return convertFromInnerType (fromList.get (index));
   }
   
   public Iterator<OE> iterator() {
      return new MyIterator (fromList.iterator());
   }

   public ListIterator<OE> listIterator() {
      return new MyListIterator (fromList.listIterator());
   }

   public ListIterator<OE> listIterator (int index) {
      return new MyListIterator (fromList.listIterator (index));
   }

   @Override
   public OE remove (int index) {
      return convertFromInnerType (fromList.remove (index));
   }

   @Override
   public OE set (int index, OE element) {
      IE from = convertToInnerType (element);
      from = fromList.set (index, from);
      return convertFromInnerType (from);
   }

   @Override
   public int size () {
      return fromList.size();
   }

   private class MyIterator implements Iterator<OE> {
      private Iterator<IE> itr;

      public MyIterator (Iterator<IE> itr) {
         this.itr = itr;
      }

      public boolean hasNext() {
         return itr.hasNext();
      }

      public OE next() {
         return convertFromInnerType (itr.next());
      }
      
      public void remove() {
         itr.remove();
      }
   }

   private class MyListIterator implements ListIterator<OE> {
      private ListIterator<IE> itr;
   
      public MyListIterator (ListIterator<IE> itr) {
         this.itr = itr;
      }

      public void add (OE o) {
         IE fe = convertToInnerType (o);
         itr.add (fe);
      }

      public boolean hasNext() {
         return itr.hasNext();
      }

      public boolean hasPrevious() {
         return itr.hasPrevious();
      }

      public OE next() {
         return convertFromInnerType (itr.next());
      }

      public int nextIndex() {
         return itr.nextIndex();
      }

      public OE previous() {
         return convertFromInnerType (itr.previous());
      }

      public int previousIndex() {
         return itr.previousIndex();
      }

      public void remove() {
         itr.remove();
      }

      public void set (OE o) {
         IE fe = convertToInnerType (o);
         itr.set (fe);
      }
   }

\end{hide}
\end{code}

\begin{code}\begin{hide}
}\end{hide}
\end{code}
