\defclass{PolInterp}

Represents a polynomial that interpolates through a set of points. More
specifically, let $(x_0,y_0), \ldots, (x_n, y_n)$ be a set of points and
$p(x)$ the constructed polynomial of degree $n$. Then, for $i=0,\ldots,n$,
$p(x_i)=y_i$.

\bigskip\hrule

\begin{code}
\begin{hide}
/*
 * Class:        PolInterp
 * Description:  polynomial that interpolates through a set of points
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.functionfit;\begin{hide}

import java.io.Serializable;
import umontreal.iro.lecuyer.functions.Polynomial;

import umontreal.iro.lecuyer.util.PrintfFormat;
import cern.colt.matrix.DoubleMatrix2D;
import cern.colt.matrix.impl.DenseDoubleMatrix2D;
import cern.colt.matrix.linalg.Algebra;
\end{hide}

public class PolInterp extends Polynomial implements Serializable\begin{hide} {
   private static final long serialVersionUID = -710451931485296501L;
   private static final Algebra alg = new Algebra ();
   private double[] x;
   private double[] y;
\end{hide}
\end{code}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection* {Constructors}
\begin{code}

   public PolInterp (double[] x, double[] y)\begin{hide} {
      super (getCoefficients (x, y));
      this.x = x.clone ();
      this.y = y.clone ();
   }\end{hide}
\end{code}
\begin{tabb}   Constructs a new polynomial interpolating through the given points
 \texttt{(x[0], y[0]), ..., (x[n], y[n])}. This constructs a polynomial of
 degree \texttt{n} from \texttt{n+1} points.
\end{tabb}
\begin{htmlonly}
   \param{x}{the $x$ coordinates of the points.}
   \param{y}{the $y$ coordinates of the points.}
   \exception{NullPointerException}{if \texttt{x} or \texttt{y} are \texttt{null}.}
   \exception{IllegalArgumentException}{if the lengths of \texttt{x} and \texttt{y} are different,
               or if less than two points are specified.}
\end{htmlonly}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection* {Methods}

\begin{code}

   public static double[] getCoefficients (double[] x, double[] y)\begin{hide} {
      if (x.length != y.length)
         throw new IllegalArgumentException (
               "x and y must have the same length");
      if (x.length <= 1)
         throw new IllegalArgumentException ("At least two points are needed");
      final DoubleMatrix2D u = new DenseDoubleMatrix2D (x.length, x.length);
      for (int i = 0; i < x.length; i++) {
         double v = 1;
         for (int j = 0; j < x.length; j++) {
            u.setQuick (i, j, v);
            v *= x[i];
         }
      }
      final DoubleMatrix2D yMat = new DenseDoubleMatrix2D (x.length, 1);
      yMat.viewColumn (0).assign (y);
      final DoubleMatrix2D bMat = alg.solve (u, yMat);
      return bMat.viewColumn (0).toArray ();
   }\end{hide}
\end{code}
\begin{tabb} Computes and returns the coefficients the polynomial interpolating
 through the given points \texttt{(x[0], y[0]), ..., (x[n], y[n])}. 
 This polynomial has degree \texttt{n} and there are \texttt{n+1} coefficients.
\end{tabb}
\begin{htmlonly}
   \param{x}{the $x$ coordinates of the points.}
   \param{y}{the $y$ coordinates of the points.}
   \return{the coefficients the interpolating polynomial.}
\end{htmlonly}
\begin{code}

   public double[] getX()\begin{hide} {
      return x.clone ();
   }\end{hide}
\end{code}
\begin{tabb}   Returns the $x$ coordinates of the interpolated points.
\end{tabb}
\begin{htmlonly}
   \return{the $x$ coordinates of the interpolated points.}
\end{htmlonly}
\begin{code}

   public double[] getY()\begin{hide} {
      return y.clone ();
   }\end{hide}
\end{code}
\begin{tabb}   Returns the $y$ coordinates of the interpolated points.
\end{tabb}
\begin{htmlonly}
   \return{the $y$ coordinates of the interpolated points.}
\end{htmlonly}
\begin{code}

   public static String toString (double[] x, double[] y)\begin{hide} {
      final StringBuilder sb = new StringBuilder ("Points: ");
      for (int i = 0; i < x.length; i++) {
         if (i > 0)
            sb.append (", ");
         final String xstr = PrintfFormat.format (8, 3, 3, x[i]);
         final String ystr = PrintfFormat.format (8, 3, 3, y[i]);
         sb.append ('(').append (xstr).append (", ").append (ystr).append (')');
      }
      return sb.toString ();
   }

   @Override
\end{hide}
\end{code}
\begin{tabb}   Makes a string representation of a set of points.
\end{tabb}
\begin{htmlonly}
   \param{x}{the $x$ coordinates of the points.}
   \param{y}{the $y$ coordinates of the points.}
   \return{the string representing the points.}
\end{htmlonly}
\begin{code}

   public String toString()\begin{hide} {
      return toString (x, y);
   }\end{hide}
\end{code}
\begin{tabb}   Calls \method{toString}~\texttt{(double[], double[])} with the
 associated  points.
\end{tabb}
\begin{htmlonly}
   \return{a string containing the points.}
\end{htmlonly}
\begin{code}\begin{hide}

   public PolInterp clone() {
      final PolInterp p = (PolInterp) super.clone ();
      p.x = x.clone ();
      p.y = y.clone ();
      return p;
   }
}\end{hide}
\end{code}
