\defclass {BoxChart}

% Extends \externalclass{umontreal.iro.lecuyer.charts}{CategoryChart}.
This class provides tools to create and manage box-and-whisker  plots.
Each \class{BoxChart} object is linked with a
\externalclass{umontreal.iro.lecuyer.charts}{BoxSeriesCollection} data set.
% It's advised to use more than 3 series.

A boxplot is a convenient way of viewing sets of numerical data through their
summaries: the smallest observation, first quartile ($Q_1 = x_{.25}$), median
 ($Q_2 = x_{.5}$), third quartile ($Q_3 = x_{.75}$), and largest observation.
Sometimes, the mean and the outliers are also plotted.

In the charts created by this class, the box has its lower limit at $Q_1$
and its upper limit at $Q_3$. The median is indicated by the line inside the box,
while the mean is at the center of the filled circle inside the box.
Define the interquartile range as ($Q_3 - Q_1$).
Any data observation which is more than $1.5(Q_3 - Q_1)$ lower than the first
quartile or $1.5(Q_3 - Q_1)$ higher than the third quartile is considered an
outlier. The smallest and the largest values that are not outliers are connected
to the box with a vertical line or "whisker" which is ended by a horizontal line.
Outliers are indicated by hollow circles outside the whiskers. Triangles
indicate the existence of very far outliers.

\bigskip\hrule

\begin{code}
\begin{hide}
/*
 * Class:        BoxChart
 * Description:  
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.charts;\begin{hide}

import   org.jfree.chart.axis.NumberAxis;
import   org.jfree.chart.ChartFactory;
import   org.jfree.chart.ChartPanel;
import   org.jfree.chart.plot.CategoryPlot;
import   org.jfree.chart.renderer.category.BoxAndWhiskerRenderer;
import   org.jfree.data.statistics.DefaultBoxAndWhiskerCategoryDataset;

import   java.util.Locale;
import   java.util.Formatter;
import   javax.swing.JFrame;
\end{hide}


public class BoxChart extends CategoryChart \begin{hide} {

   protected void init (String title, String XLabel, String YLabel) {
      // create the chart...
      chart = ChartFactory.createBoxAndWhiskerChart (
         title,                         // chart title
         XLabel,                        // x axis label
         YLabel,                        // y axis label
         (DefaultBoxAndWhiskerCategoryDataset)dataset.getSeriesCollection(), // data
         true                          // include legend
      );

      ((CategoryPlot)chart.getPlot()).setRenderer(dataset.getRenderer());
      // Initialize axis variables
      initAxis();
   }

   protected void initAxis(){
      YAxis = new Axis((NumberAxis)((CategoryPlot) chart.getPlot()).getRangeAxis(),
                        Axis.ORIENTATION_VERTICAL);
      setAutoRange();
   }
\end{hide}
\end{code}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Constructors}

\begin{code}

   public BoxChart() \begin{hide} {
      super();
      dataset = new BoxSeriesCollection();
      init (null, null, null);
   }\end{hide}
\end{code}
\begin{tabb}
   Initializes a new \texttt{BoxChart} instance with an empty data set.
\end{tabb}
\begin{code}

   public BoxChart (String title, String XLabel, String YLabel,
                    double[] data, int numPoints) \begin{hide} {
      super();
      dataset = new BoxSeriesCollection(data, numPoints);
      init (title, XLabel, YLabel);
   }\end{hide}
\end{code}
\begin{tabb}
Initializes a new \texttt{BoxChart} instance with data \texttt{data}.
\texttt{title} is a title, \texttt{XLabel} is a short description of the
$x$-axis, and \texttt{YLabel} a short description of the $y$-axis.
The input parameter \texttt{data} represents a set of plotting data.
 Only \emph{the first} \texttt{numPoints} of \texttt{data} will
  be considered for the plot.
\end{tabb}
\begin{htmlonly}
   \param{title}{chart title.}
   \param{XLabel}{Label on $x$-axis.}
   \param{YLabel}{Label on $y$-axis.}
   \param{data}{point sets.}
   \param{numPoints}{Number of points to plot}
\end{htmlonly}
\begin{code}

   public BoxChart (String title, String XLabel, String YLabel,
                    double[]... data) \begin{hide} {
      super();
      dataset = new BoxSeriesCollection(data);
      init (title, XLabel, YLabel);
   }\end{hide}
\end{code}
\begin{tabb}
Initializes a new \texttt{BoxChart} instance with data \texttt{data}.
\texttt{title} sets a title, \texttt{XLabel} is a short description of the
$x$-axis, and \texttt{YLabel} is a short description of the $y$-axis.
The input parameter \texttt{data} represents a set of plotting data.
\end{tabb}
\begin{htmlonly}
   \param{title}{chart title.}
   \param{XLabel}{Label on $x$-axis.}
   \param{YLabel}{Label on $y$-axis.}
   \param{data}{series of point sets.}
\end{htmlonly}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Methods}

\begin{code}

   public int add (double[] data) \begin{hide} {
      return add(data, data.length);
   }\end{hide}
\end{code}
\begin{tabb}
   Adds a data series into the series collection. Vector \texttt{data} represents
   a set of plotting data.
\end{tabb}
\begin{htmlonly}
   \param{data}{point sets.}
   \return{Integer that represent the new point set's position in the JFreeChart \texttt{BoxSeriesCollection} object.}
\end{htmlonly}
\begin{code}

   public int add (double[] data, int numPoints) \begin{hide} {
      int seriesIndex = getSeriesCollection().add(data, numPoints);
      initAxis();
      return seriesIndex;
   }\end{hide}
\end{code}
\begin{tabb}
   Adds a data series into the series collection. Vector \texttt{data} represents
   a set of plotting data. Only \emph{the first} \texttt{numPoints} of
   \texttt{data} will be taken into account for the new series.
\end{tabb}
\begin{htmlonly}
   \param{data}{point set.}
   \param{numPoints}{number of points to add.}
   \return{Integer that represent the new point set's position in the JFreeChart \texttt{BoxSeriesCollection} object.}
\end{htmlonly}
\begin{code}

   public BoxSeriesCollection getSeriesCollection() \begin{hide} {
      return (BoxSeriesCollection)dataset;
   }\end{hide}
\end{code}
\begin{tabb}
   Returns the chart's dataset.
\end{tabb}
\begin{htmlonly}
   \return{the chart's dataset.}
\end{htmlonly}
\begin{code}

   public void setSeriesCollection (BoxSeriesCollection dataset) \begin{hide} {
      this.dataset = dataset;
   }\end{hide}
\end{code}
\begin{tabb}
   Links a new dataset to the current chart.
\end{tabb}
\begin{htmlonly}
   \param{dataset}{new dataset.}
\end{htmlonly}
\begin{code}

   public void setFillBox (boolean fill) \begin{hide} {
      ((BoxAndWhiskerRenderer)dataset.getRenderer()).setFillBox(fill);
   }\end{hide}
\end{code}
\begin{tabb}
   Sets \texttt{fill} to \texttt{true}, if the boxes are to be filled.
\end{tabb}
\begin{htmlonly}
   \param{fill}{true if the boxes are filled}
\end{htmlonly}
\begin{code}

   public JFrame view (int width, int height) \begin{hide} {
      JFrame myFrame;
      if(chart.getTitle() != null)
         myFrame = new JFrame ("BoxChart from SSJ : " + chart.getTitle().getText());
      else
         myFrame = new JFrame ("BoxChart from SSJ");
      ChartPanel chartPanel = new ChartPanel (chart);
      chartPanel.setPreferredSize (new java.awt.Dimension(width, height));
      myFrame.setContentPane (chartPanel);
      myFrame.pack ();
      myFrame.setDefaultCloseOperation (JFrame.DISPOSE_ON_CLOSE);
      myFrame.setLocationRelativeTo (null);
      myFrame.setVisible (true);
      return myFrame;
   }\end{hide}
\end{code}
\begin{tabb}
   Displays chart on the screen using Swing.
   This method creates an application containing a chart panel displaying
   the chart.  The created frame is positioned on-screen, and displayed before
   it is returned. The circle represents the mean, the dark line inside the box
   is the median, the box limits are the first and third quartiles,
   the lower whisker (the lower line outside the box) is the first decile,
   and the upper whisker is the ninth decile. The outliers, if any, are
   represented by empty circles, or arrows if outside the range bounds.
\end{tabb}
\begin{htmlonly}
   \param{width}{frame width.}
   \param{height}{frame height.}
   \return{frame containing the chart.};
\end{htmlonly}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Latex-specific method}

\begin{code}

  public String toLatex (double width, double height) \begin{hide} {
  throw new UnsupportedOperationException(" NOT implemented yet");
/*
      double yunit;
      double[] save = new double[4];

      if(dataset.getSeriesCollection().getColumnCount() == 0)
         throw new IllegalArgumentException("Empty chart");

      //Calcul des parametres d'echelle et de decalage
      double YScale = computeYScale(YAxis.getTwinAxisPosition());


      yunit = height / ( (Math.max(YAxis.getAxis().getRange().getUpperBound(), YAxis.getTwinAxisPosition()) * YScale) - (Math.min(YAxis.getAxis().getRange().getLowerBound(), YAxis.getTwinAxisPosition()) * YScale) );
      //taille d'une unite en y et en cm dans l'objet "tikzpicture"

      Formatter formatter = new Formatter(Locale.US);

      // Entete du document
      formatter.format("\\documentclass[12pt]{article}%n%n");
      formatter.format("\\usepackage{tikz}%n\\usetikzlibrary{plotmarks}%n\\begin{document}%n%n");
      if(chart.getTitle() != null)
         formatter.format("%% PGF/TikZ picture from SSJ : %s%n", chart.getTitle().getText());
      else
         formatter.format("%% PGF/TikZ picture from SSJ %n");
      formatter.format("%%  YScale = %s, YShift = %s%n", YScale,  YAxis.getTwinAxisPosition());
      formatter.format("%%        and thisFileYValue = (originalSeriesYValue+YShift)*YScale%n%n");
      if (chart.getTitle() != null)
         formatter.format("\\begin{figure}%n");
      formatter.format("\\begin{center}%n");
      formatter.format("\\begin{tikzpicture}[y=%scm]%n", yunit);
      formatter.format("\\footnotesize%n");
      if(grid)
         formatter.format("\\draw[color=lightgray] (%s) grid[ystep=%s] (%s);%n",
            (Math.min(YAxis.getAxis().getRange().getLowerBound(), YAxis.getTwinAxisPosition())-YAxis.getTwinAxisPosition()) * YScale,
            ystepGrid*YScale,
            (Math.max(YAxis.getAxis().getRange().getUpperBound(), YAxis.getTwinAxisPosition())-YAxis.getTwinAxisPosition()) * YScale );
      formatter.format("%s", YAxis.toLatex(YScale) );

      formatter.format("%s", dataset.toLatex(YScale, YAxis.getTwinAxisPosition(),      YAxis.getAxis().getLowerBound(), YAxis.getAxis().getUpperBound()));

      formatter.format("\\end{tikzpicture}%n");
      formatter.format("\\end{center}%n");
      if (chart.getTitle() != null) {
         formatter.format("\\caption{");
         formatter.format(chart.getTitle().getText());
         formatter.format("}%n\\end{figure}%n");
      }
      formatter.format("\\end{document}%n");
      return formatter.toString();
*/
   }\end{hide}
\end{code}
\begin{tabb}
NOT IMPLEMENTED.
%   Exports the chart to a \LaTeX\ source code using PGF/TikZ.
%   This method constructs and returns a string that can be written to
%   a \LaTeX\ document to render the plot. \texttt{width} and \texttt{height}
%   represents the width and the height of the produced chart. These dimensions
%   do not take into account the axes and labels extra space.
\end{tabb}
\begin{htmlonly}
   \param{width}{Chart's width in centimeters.}
   \param{height}{Chart's height in centimeters.}
   \return{LaTeX source code.}
\end{htmlonly}
\begin{code}
\begin{hide}
}\end{hide}
\end{code}
