\defclass {EmpiricalChart}

% Extends \externalclass{umontreal.iro.lecuyer}{Chart}.
This class provides additional tools to create and manage empirical
plots. Empirical plots are used to plot empirical distributions. The
\class{EmpiricalChart} class is the simplest way to produce empirical
  plots only. Each \class{EmpiricalChart} object is linked with an
\externalclass{umontreal.iro.lecuyer.charts}{EmpiricalSeriesCollection}
data set.

\bigskip\hrule
\begin{code}
\begin{hide}
/*
 * Class:        EmpiricalChart
 * Description:  
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.charts;\begin{hide}

import   umontreal.iro.lecuyer.stat.TallyStore;

import   org.jfree.chart.ChartPanel;
import   org.jfree.chart.ChartFactory;
import   org.jfree.chart.axis.NumberAxis;
import   org.jfree.chart.plot.XYPlot;
import   org.jfree.chart.plot.PlotOrientation;
import   org.jfree.data.xy.XYSeriesCollection;
import   org.jfree.data.xy.XYSeries;

import   cern.colt.list.DoubleArrayList;

import   java.util.ListIterator;
import   java.util.Locale;
import   java.util.Formatter;
import   javax.swing.JFrame;\end{hide}

public class EmpiricalChart extends XYChart \begin{hide} {

   protected void init (String title, String XLabel, String YLabel) {
      // create the chart...
      chart = ChartFactory.createXYLineChart(
         title,                    // chart title
         XLabel,                   // x axis label
         YLabel,                   // y axis label
         dataset.getSeriesCollection(), // data
         PlotOrientation.VERTICAL,
         true,                     // include legend
         true,                     // tool tips
         false                     // urls
      );
      ((XYPlot)chart.getPlot()).setRenderer(dataset.getRenderer());
      // Initialize axis variables
      XAxis = new Axis((NumberAxis)((XYPlot) chart.getPlot()).getDomainAxis(),
                Axis.ORIENTATION_HORIZONTAL);
      YAxis = new Axis((NumberAxis)((XYPlot) chart.getPlot()).getRangeAxis(),
                Axis.ORIENTATION_VERTICAL);
      fixZeroPoint();
   }


   private void fixZeroPoint() {
      // reset the first point (x0, 0) with x0 at the beginning of x-axis
      double xmin = Math.min (XAxis.getAxis().getRange().getLowerBound(),
                                                   XAxis.getTwinAxisPosition());
      XYSeriesCollection col = (XYSeriesCollection)dataset.getSeriesCollection();
      for (int i = 0; i < col.getSeriesCount(); i++) {
         XYSeries ser = col.getSeries (i);
         ser.remove(0);   // remove temporary 0-point
         ser.add(xmin, 0); // replace
      }
   }\end{hide}
\end{code}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Constructors}


\begin{code}
   public EmpiricalChart() \begin{hide} {
      super();
      dataset = new EmpiricalSeriesCollection();
      init (null, null, null);
   }\end{hide}
\end{code}
\begin{tabb}
   Initializes a new \texttt{EmpiricalChart} instance with an empty data set.
\end{tabb}
\begin{code}

   public EmpiricalChart (String title, String XLabel, String YLabel,
                          double[]... data) \begin{hide} {
      super();
      dataset = new EmpiricalSeriesCollection(data);
      init (title, XLabel, YLabel);
   }\end{hide}
\end{code}
\begin{tabb}
   Initializes a new \texttt{EmpiricalChart} instance with data \texttt{data}.
   \texttt{title} is a title, \texttt{XLabel} is a short description of
   the $x$-axis and \texttt{YLabel} a short description of the $y$-axis.
   The input vectors \texttt{data} represents a collection of observation sets.
   Each vector of \texttt{data} represents a $x$-coordinates set.
   Therefore \texttt{data}$[i], i = 0,\ldots,n-1$, is used to draw the $i$-th plot.
   The values of each observation set \texttt{data}$[i]$ \emph{must be sorted}
   in increasing order.
\end{tabb}
\begin{htmlonly}
   \param{title}{chart title.}
   \param{XLabel}{Label on $x$-axis.}
   \param{YLabel}{Label on $y$-axis.}
   \param{data}{series of point sets.}
\end{htmlonly}
\begin{code}

   public EmpiricalChart (String title, String XLabel, String YLabel,
                          double[] data, int numPoints) \begin{hide} {
      super();
      dataset = new EmpiricalSeriesCollection(data, numPoints);
      init (title, XLabel, YLabel);
   }\end{hide}
\end{code}
\begin{tabb}
   Initializes a new \texttt{EmpiricalChart} instance with a set of points
   \texttt{data}. \texttt{title} is a title, \texttt{XLabel} is a short
   description of the $x$-axis and \texttt{YLabel} a short description of the
   $y$-axis. Vector \texttt{data} represents a $x$-coordinates set. The values
   of this observation set \emph{must be sorted} in increasing order. Only
   \emph{the first} \texttt{numPoints} of \texttt{data} will
   be considered to plot.
\end{tabb}
\begin{htmlonly}
   \param{title}{chart title.}
   \param{XLabel}{Label on $x$-axis.}
   \param{YLabel}{Label on $y$-axis.}
   \param{data}{series of point sets.}
   \param{numPoints}{number of points to plot}
\end{htmlonly}
\begin{code}

   public EmpiricalChart (String title, String XLabel, String YLabel,
                          DoubleArrayList... data) \begin{hide} {
      super();
      dataset = new EmpiricalSeriesCollection(data);
      init (title, XLabel, YLabel);
   }\end{hide}
\end{code}
\begin{tabb}
   Similar to the above constructor, but with \texttt{DoubleArrayList}.
   A \externalclass{cern.colt.list}{DoubleArrayList} from the Colt library is
   used to store data. The values of each observation set \texttt{data}$[i]$
   \emph{must be sorted} in increasing order.
\end{tabb}
\begin{htmlonly}
   \param{title}{chart title.}
   \param{XLabel}{Label on $x$-axis.}
   \param{YLabel}{Label on $y$-axis.}
   \param{data}{series of point sets.}
\end{htmlonly}
\begin{code}

   public EmpiricalChart (String title, String XLabel, String YLabel,
                          TallyStore... tallies) \begin{hide} {
      super();
      dataset = new EmpiricalSeriesCollection(tallies);
      init (title, XLabel, YLabel);
   }\end{hide}
\end{code}
\begin{tabb}
   Initializes a new \texttt{EmpiricalChart} instance with data arrays
   contained in each
   \externalclass{umontreal.iro.lecuyer.stat}{TallyStore} object. The input
   parameter \texttt{tallies} represents a collection of observation sets.
   Therefore, the $i$-th \texttt{tallies} is used to draw
     the $i$-th plot.
\end{tabb}
\begin{htmlonly}
   \param{title}{chart title.}
   \param{XLabel}{Label on $x$-axis.}
   \param{YLabel}{Label on $y$-axis.}
   \param{tallies}{series of observation sets.}
\end{htmlonly}
\begin{code}

   public EmpiricalChart (String title, String XLabel, String YLabel,
                          XYSeriesCollection data) \begin{hide} {
      super();
      dataset = new EmpiricalSeriesCollection(data);
      init (title, XLabel, YLabel);
   }\end{hide}
\end{code}
\begin{tabb}
   Initializes a new \texttt{EmpiricalChart} instance with data \texttt{data}.
   The input parameter \texttt{data} represents a set of plotting data.
   \texttt{XYSeriesCollection} is a \texttt{JFreeChart}-like container class
    used to store and manage observation sets.
\end{tabb}
\begin{htmlonly}
   \param{title}{chart title.}
   \param{XLabel}{Label on $x$-axis.}
   \param{YLabel}{Label on $y$-axis.}
   \param{data}{series collection.}
\end{htmlonly}


%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Methods}

\begin{code}

   public EmpiricalSeriesCollection getSeriesCollection() \begin{hide} {
      return (EmpiricalSeriesCollection)dataset;
   }\end{hide}
\end{code}
\begin{tabb}
   Returns the chart's dataset.
\end{tabb}
\begin{htmlonly}
   \return{the chart's dataset.}
\end{htmlonly}
\begin{code}

   public void setSeriesCollection (EmpiricalSeriesCollection dataset) \begin{hide} {
      this.dataset = dataset;
   }\end{hide}
\end{code}
\begin{tabb}
   Links a new dataset to the current chart.
\end{tabb}
\begin{htmlonly}
   \param{dataset}{new dataset.}
\end{htmlonly}
\begin{code}

   public void setTicksSynchro (int s) \begin{hide} {
      XYSeriesCollection seriesCollection = (XYSeriesCollection)this.dataset.getSeriesCollection();
      double[] values = new double[seriesCollection.getItemCount(s)];

      for(int i = 0; i < seriesCollection.getItemCount(s); i++)
         values[i] = seriesCollection.getXValue(s, i);

      XAxis.setLabels(values);
   }\end{hide}
\end{code}
\begin{tabb}
   Synchronizes $x$-axis ticks to the $s$-th series $x$-values.
\end{tabb}
\begin{htmlonly}
   \param{s}{series used to define ticks.}
\end{htmlonly}
\begin{code}

   public JFrame view (int width, int height) \begin{hide} {
      JFrame myFrame;
      if(chart.getTitle() != null)
         myFrame = new JFrame("EmpiricalChart from SSJ : " + chart.getTitle().getText());
      else
         myFrame = new JFrame("EmpiricalChart from SSJ");
      ChartPanel chartPanel = new ChartPanel(chart);
      chartPanel.setPreferredSize(new java.awt.Dimension(width, height));
      myFrame.setContentPane(chartPanel);
      myFrame.pack();
      myFrame.setDefaultCloseOperation (JFrame.DISPOSE_ON_CLOSE);
      myFrame.setLocationRelativeTo (null);
      myFrame.setVisible(true);
      return myFrame;
   }\end{hide}
\end{code}
\begin{tabb}
   Displays chart on the screen using Swing.
   This method creates an application containing a chart panel displaying
   the chart.  The created frame is positioned on-screen, and displayed before
   it is returned. The \texttt{width} and the \texttt{height}
   of the chart are measured in pixels.
\end{tabb}
\begin{htmlonly}
   \param{width}{frame width in pixels.}
   \param{height}{frame height in pixels.}
   \return{frame containing the chart.}
\end{htmlonly}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{\LaTeX-specific method}

\begin{code}

   public String toLatex (double width, double height) \begin{hide} {
      double xunit, yunit;
      double[] save = new double[4];

      if(dataset.getSeriesCollection().getSeriesCount() == 0)
         throw new IllegalArgumentException("Empty chart");

      //Calcul des parametres d'echelle et de decalage
      double XScale = computeXScale(XAxis.getTwinAxisPosition());
      double YScale = computeYScale(YAxis.getTwinAxisPosition());

      xunit = width / ((Math.max(XAxis.getAxis().getRange().getUpperBound(), XAxis.getTwinAxisPosition()) * XScale) - (Math.min(XAxis.getAxis().getRange().getLowerBound(), XAxis.getTwinAxisPosition()) * XScale));
      //taille d'une unite en x et en cm dans l'objet "tikzpicture"
      yunit = height / ((Math.max(YAxis.getAxis().getRange().getUpperBound(), YAxis.getTwinAxisPosition()) * YScale) - (Math.min(YAxis.getAxis().getRange().getLowerBound(), YAxis.getTwinAxisPosition()) * YScale));
      //taille d'une unite en y et en cm dans l'objet "tikzpicture"

      Formatter formatter = new Formatter(Locale.US);

      /*Entete du document*/
      if (latexDocFlag) {
         formatter.format("\\documentclass[12pt]{article}%n%n");
         formatter.format("\\usepackage{tikz}%n\\usetikzlibrary{plotmarks}%n\\begin{document}%n%n");
      }
      if(chart.getTitle() != null)
         formatter.format("%% PGF/TikZ picture from SSJ : %s%n", chart.getTitle().getText());
      else
         formatter.format("%% PGF/TikZ picture from SSJ %n");
      formatter.format("%% XScale = %s,  YScale = %s,  XShift = %s,  YShift = %s%n", XScale, YScale, XAxis.getTwinAxisPosition(), YAxis.getTwinAxisPosition());
      formatter.format("%% Therefore, thisFileXValue = (originalSeriesXValue+XShift)*XScale%n");
      formatter.format("%%        and thisFileYValue = (originalSeriesYValue+YShift)*YScale%n%n");
      if (chart.getTitle() != null)
         formatter.format("\\begin{figure}%n");
      formatter.format("\\begin{center}%n");
      formatter.format("\\begin{tikzpicture}[x=%scm, y=%scm]%n", xunit, yunit);
      formatter.format("\\footnotesize%n");
      if (grid)
         formatter.format("\\draw[color=lightgray] (%s, %s) grid[xstep = %s, ystep=%s] (%s, %s);%n",
            (Math.min(XAxis.getAxis().getRange().getLowerBound(), XAxis.getTwinAxisPosition())-XAxis.getTwinAxisPosition()) * XScale,
            (Math.min(YAxis.getAxis().getRange().getLowerBound(), YAxis.getTwinAxisPosition())-YAxis.getTwinAxisPosition()) * YScale,
            xstepGrid*XScale, ystepGrid*YScale,
            (Math.max(XAxis.getAxis().getRange().getUpperBound(), XAxis.getTwinAxisPosition())-XAxis.getTwinAxisPosition()) * XScale,
            (Math.max(YAxis.getAxis().getRange().getUpperBound(), YAxis.getTwinAxisPosition())-YAxis.getTwinAxisPosition()) * YScale);
      setTick0Flags();
      formatter.format("%s", XAxis.toLatex(XScale));
      formatter.format("%s", YAxis.toLatex(YScale));

      formatter.format("%s", dataset.toLatex(XScale, YScale, XAxis.getTwinAxisPosition(), YAxis.getTwinAxisPosition(),
                                                            XAxis.getAxis().getLowerBound(), XAxis.getAxis().getUpperBound(),
                                                            YAxis.getAxis().getLowerBound(), YAxis.getAxis().getUpperBound()));

      formatter.format("\\end{tikzpicture}%n");
      formatter.format("\\end{center}%n");
      if (chart.getTitle() != null) {
         formatter.format("\\caption{");
         formatter.format(chart.getTitle().getText());
         formatter.format("}%n\\end{figure}%n");
      }
      if (latexDocFlag)
         formatter.format("\\end{document}%n");
      return formatter.toString();
   }\end{hide}
\end{code}

\begin{code}
\begin{hide}
}\end{hide}
\end{code}
