\defclass {MultipleDatasetChart}

Provides tools to plot many datasets on the same chart.
This class is mainly used to draw plots with different styles.
Class \externalclass{umontreal.iro.lecuyer.charts}{XYChart} and
its subclasses are to be preferred to draw simple charts with one style.
Datasets are stored in an \texttt{ArrayList}. The first dataset is called
as the \textit{primary dataset}.

\bigskip\hrule
\begin{code}
\begin{hide}
/*
 * Class:        MultipleDatasetChart
 * Description:  
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.charts;\begin{hide}

import   org.jfree.chart.JFreeChart;
import   org.jfree.chart.ChartPanel;
import   org.jfree.chart.ChartFactory;
import   org.jfree.chart.axis.NumberAxis;
import   org.jfree.chart.plot.XYPlot;
import   org.jfree.chart.plot.PlotOrientation;

import   java.util.Locale;
import   java.util.Formatter;
import   java.util.ArrayList;
import   javax.swing.JFrame;\end{hide}

public class MultipleDatasetChart \begin{hide} {

   protected ArrayList<SSJXYSeriesCollection> datasetList;
   protected Axis XAxis;
   protected Axis YAxis;
   protected JFreeChart chart;
   protected boolean latexDocFlag = true;

   protected boolean autoRange = true;
   protected double[] manualRange;

   protected boolean grid = false;
   protected double xstepGrid;
   protected double ystepGrid;

\end{hide}
\end{code}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Constructors}

\begin{code}
   public MultipleDatasetChart() \begin{hide} {
      super();

      // create the chart...
      chart = ChartFactory.createXYLineChart(
         null,                     // chart title
         null,                     // x axis label
         null,                     // y axis label
         null, // data
         PlotOrientation.VERTICAL,
         true,                     // include legend
         true,                     // tool tips
         false                     // urls
      );

      datasetList = new ArrayList<SSJXYSeriesCollection>();
      // Initialize axis variables
      XAxis = new Axis((NumberAxis)((XYPlot) chart.getPlot()).getDomainAxis(),
                 Axis.ORIENTATION_HORIZONTAL);
      YAxis = new Axis((NumberAxis)((XYPlot) chart.getPlot()).getRangeAxis(),
                 Axis.ORIENTATION_VERTICAL);
   }\end{hide}
\end{code}
\begin{tabb}
   Initializes a new \texttt{MultipleDatasetChart}.
\end{tabb}
\begin{code}

   public MultipleDatasetChart (String title, String XLabel, String YLabel) \begin{hide} {
      // create the chart...
      chart = ChartFactory.createXYLineChart(
         title,                    // chart title
         XLabel,                   // x axis label
         YLabel,                   // y axis label
         null,                     // data
         PlotOrientation.VERTICAL,
         true,                     // include legend
         true,                     // tool tips
         false                     // urls
      );

      datasetList = new ArrayList<SSJXYSeriesCollection>();
      //Initialize axis variables
      XAxis = new Axis((NumberAxis)((XYPlot) chart.getPlot()).getDomainAxis(),
              Axis.ORIENTATION_HORIZONTAL);
      YAxis = new Axis((NumberAxis)((XYPlot) chart.getPlot()).getRangeAxis(),
              Axis.ORIENTATION_VERTICAL);
   }\end{hide}
\end{code}
\begin{tabb}
   Initializes a new \texttt{MultipleDatasetChart} instance.
   \texttt{title} sets a title, \texttt{XLabel} is a short description of
   the $x$-axis, and \texttt{YLabel} is a short description of the $y$-axis.
\end{tabb}
\begin{htmlonly}
   \param{title}{chart title.}
   \param{XLabel}{Label on $x$-axis.}
   \param{YLabel}{Label on $y$-axis.}
\end{htmlonly}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Methods}

\begin{code}

   public JFreeChart getJFreeChart() \begin{hide} {
      return chart;
   }\end{hide}
\end{code}
\begin{tabb}
   Returns the \texttt{JFreeChart} variable associated with this chart.
\end{tabb}
\begin{htmlonly}
   \return{the associated JFreeChart variable.}
\end{htmlonly}
\begin{code}

   public Axis getXAxis() \begin{hide} {
      return XAxis;
   }\end{hide}
\end{code}
\begin{tabb}
   Returns the chart's domain axis ($x$-axis) object.
\end{tabb}
\begin{htmlonly}
   \return{chart's domain axis ($x$-axis) object.}
\end{htmlonly}
\begin{code}

   public Axis getYAxis() \begin{hide} {
      return YAxis;
   }\end{hide}
\end{code}
\begin{tabb}
   Returns the chart's range axis ($y$-axis) object.
\end{tabb}
\begin{htmlonly}
   \return{chart's range axis ($y$-axis) object.}
\end{htmlonly}
\begin{code}

   public String getTitle() \begin{hide} {
      return chart.getTitle().getText();
   }\end{hide}
\end{code}
\begin{tabb}
   Gets the current chart title.
\end{tabb}
\begin{htmlonly}
   \return{Chart title.}
\end{htmlonly}
\begin{code}

   public void setTitle (String title) \begin{hide} {
      chart.setTitle(title);
   }\end{hide}
\end{code}
\begin{tabb}
   Sets a title to the chart. This title will appear on the chart displayed by method \method{view}{}.
\end{tabb}
\begin{htmlonly}
   \param{title}{chart title.}
\end{htmlonly}
\begin{code}

   public void setAutoRange() \begin{hide} {
      autoRange = true;
      double[][] temp = new double[2][datasetList.size()];
      for(int i = 0; i<datasetList.size(); i++) {
         temp[0][i] = (datasetList.get(i).getDomainBounds())[0];
         temp[1][i] = (datasetList.get(i).getDomainBounds())[1];
      }
      XAxis.getAxis().setLowerBound(min(temp[0]));
      XAxis.getAxis().setUpperBound(max(temp[1]));

      for(int i = 0; i<datasetList.size(); i++) {
         temp[0][i] = (datasetList.get(i).getRangeBounds())[0];
         temp[1][i] = (datasetList.get(i).getRangeBounds())[1];
      }
      YAxis.getAxis().setLowerBound(min(temp[0]));
      YAxis.getAxis().setUpperBound(max(temp[1]));
   }\end{hide}
\end{code}
\begin{tabb}
   Sets chart range to automatic values.
\end{tabb}
\begin{code}

   public void setManualRange (double[] axisRange) \begin{hide} {
      if(axisRange.length != 4)
         throw new IllegalArgumentException("axisRange must share the format: [xmin, xmax, ymin, ymax]");
      autoRange = false;
      XAxis.getAxis().setLowerBound(axisRange[0]);
      XAxis.getAxis().setUpperBound(axisRange[1]);
      YAxis.getAxis().setLowerBound(axisRange[2]);
      YAxis.getAxis().setUpperBound(axisRange[3]);
   }\end{hide}
\end{code}
\begin{tabb}
   Sets new $x$-axis and $y$-axis bounds, with format: \texttt{axisRange} = [xmin, xmax, ymin, ymax].
\end{tabb}
\begin{htmlonly}
   \param{axisRange}{new axis ranges.}
\end{htmlonly}
\begin{code}

   public int add (SSJXYSeriesCollection dataset) \begin{hide} {
      ((XYPlot)chart.getPlot()).setDataset(datasetList.size(),
                  dataset.getSeriesCollection());
      ((XYPlot)chart.getPlot()).setRenderer(datasetList.size(),
                  dataset.getRenderer());
      datasetList.add(dataset);
      if(datasetList.size() == 1) {
         XAxis.setLabelsAuto();
         YAxis.setLabelsAuto();
      }
      return datasetList.size()-1;
   }\end{hide}
\end{code}
\begin{tabb}
   Adds a new dataset to the chart at the end of the list and returns its position.
\end{tabb}
\begin{htmlonly}
   \param{dataset}{dataset to add.}
   \return{the dataset position in the list.}
\end{htmlonly}
\begin{code}

   public SSJXYSeriesCollection get() \begin{hide} {
      return datasetList.get(0);
   }\end{hide}
\end{code}
\begin{tabb}
   Gets the primary dataset.
\end{tabb}
\begin{htmlonly}
   \return{dataset.}
\end{htmlonly}
\begin{code}

   public void set (SSJXYSeriesCollection dataset) \begin{hide} {
      ((XYPlot)chart.getPlot()).setDataset(dataset.getSeriesCollection());
      ((XYPlot)chart.getPlot()).setRenderer(dataset.getRenderer());
      datasetList.set(0, dataset);
   }\end{hide}
\end{code}
\begin{tabb}
    Sets the primary dataset for the plot, replacing the existing dataset if there is one.
\end{tabb}
\begin{htmlonly}
   \param{dataset}{the new primary dataset.}
\end{htmlonly}
\begin{code}

   public SSJXYSeriesCollection get (int datasetNum) \begin{hide} {
      return datasetList.get(datasetNum);
   }\end{hide}
\end{code}
\begin{tabb}
   Gets the element at the specified position in the dataset list.
\end{tabb}
\begin{htmlonly}
   \param{datasetNum}{position in the dataset list.}
   \return{dataset.}
\end{htmlonly}
\begin{code}

   public void set (int datasetNum, SSJXYSeriesCollection dataset) \begin{hide} {
      ((XYPlot)chart.getPlot()).setDataset(datasetNum, dataset.getSeriesCollection());
      ((XYPlot)chart.getPlot()).setRenderer(datasetNum, dataset.getRenderer());
      datasetList.add(datasetNum, dataset);
   }\end{hide}
\end{code}
\begin{tabb}
   Replaces the element at the specified position in the dataset list with the specified element.
\end{tabb}
\begin{htmlonly}
   \param{datasetNum}{position in the dataset list.}
   \param{dataset}{dataset list.}
\end{htmlonly}
\begin{code}

   public ArrayList<SSJXYSeriesCollection> getList() \begin{hide} {
      return datasetList;
   }\end{hide}
\end{code}
\begin{tabb}
   Returns the dataset list.
\end{tabb}
\begin{htmlonly}
   \return{dataset list.}
\end{htmlonly}
\begin{code}

   public JFrame view (int width, int height) \begin{hide} {
      JFrame myFrame;
      if (chart.getTitle() != null)
         myFrame = new JFrame("MultipleDatasetChart from SSJ: " +
                               chart.getTitle().getText());
      else
         myFrame = new JFrame("MultipleDatasetChart from SSJ");
      ChartPanel chartPanel = new ChartPanel(chart);
      chartPanel.setPreferredSize(new java.awt.Dimension(width, height));
      myFrame.setContentPane(chartPanel);
      myFrame.pack();
      myFrame.setDefaultCloseOperation (JFrame.DISPOSE_ON_CLOSE);
      myFrame.setLocationRelativeTo (null);
      myFrame.setVisible(true);
      return myFrame;
   }\end{hide}
\end{code}
\begin{tabb}
   Displays chart on the screen using Swing.
   This method creates an application containing a chart panel displaying
   the chart. The created frame is positioned on-screen, and displayed before
   it is returned. The \texttt{width} and the \texttt{height}
   of the chart are measured in pixels.
\end{tabb}
\begin{htmlonly}
   \param{width}{frame width in pixels.}
   \param{height}{frame height in pixels.}
\end{htmlonly}
\begin{code}

   public void enableGrid (double xstep, double ystep) \begin{hide} {
      this.grid = true;
      this.xstepGrid = xstep;
      this.ystepGrid = ystep;
   }\end{hide}
\end{code}
\begin{tabb}
   Puts grid on the background. It is important to note that the grid is
always placed in such a way that it contains the axes. Thus, the grid does
   not always have an intersection at the corner points; this occurs
   only if the corner points are multiples of the stepping. \texttt{xstep}
   and \texttt{ystep} sets the stepping in each direction.
\end{tabb}
\begin{htmlonly}
   \param{xstep}{sets the stepping in the $x$-direction.}
   \param{ystep}{sets the stepping in the $y$-direction.}
\end{htmlonly}
\begin{code}

   public void disableGrid () \begin{hide} {
      this.grid = false;
   }\end{hide}
\end{code}
\begin{tabb}
   Disables the background grid.
\end{tabb}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{\LaTeX-specific method}

\begin{code}

   public String toLatex (double width, double height) \begin{hide} {
      double xunit, yunit;
      double[] save = new double[4];

      if(datasetList.size() == 0)
         throw new IllegalArgumentException("Empty chart");

      //Calcul des parametres d'echelle et de decalage
      double XScale = computeXScale(XAxis.getTwinAxisPosition());
      double YScale = computeYScale(YAxis.getTwinAxisPosition());

      xunit = width / ( (Math.max(XAxis.getAxis().getRange().getUpperBound(), XAxis.getTwinAxisPosition()) * XScale) - (Math.min(XAxis.getAxis().getRange().getLowerBound(), XAxis.getTwinAxisPosition()) * XScale) );
      //taille d'une unite en x et en cm dans l'objet "tikzpicture"
      yunit = height / ( (Math.max(YAxis.getAxis().getRange().getUpperBound(), YAxis.getTwinAxisPosition()) * YScale) - (Math.min(YAxis.getAxis().getRange().getLowerBound(), YAxis.getTwinAxisPosition()) * YScale) );
      //taille d'une unite en y et en cm dans l'objet "tikzpicture"

      Formatter formatter = new Formatter(Locale.US);

      /*Entete du document*/
      if (latexDocFlag) {
         formatter.format("\\documentclass[12pt]{article}%n%n");
         formatter.format("\\usepackage{tikz}%n\\usetikzlibrary{plotmarks}%n\\begin{document}%n%n");
      }
      if(chart.getTitle() != null)
         formatter.format("%% PGF/TikZ picture from SSJ : %s%n", chart.getTitle().getText());
      else
         formatter.format("%% PGF/TikZ picture from SSJ %n");
      formatter.format("%% XScale = %s,  YScale = %s,  XShift = %s,  YShift = %s%n", XScale, YScale, XAxis.getTwinAxisPosition(), YAxis.getTwinAxisPosition());
      formatter.format("%% Therefore, thisFileXValue = (originalSeriesXValue+XShift)*XScale%n");
      formatter.format("%%        and thisFileYValue = (originalSeriesYValue+YShift)*YScale%n%n");
      if (chart.getTitle() != null)
         formatter.format("\\begin{figure}%n");
      formatter.format("\\begin{center}%n");
      formatter.format("\\begin{tikzpicture}[x=%scm, y=%scm]%n", xunit, yunit);
      formatter.format("\\footnotesize%n");
      if(grid)
         formatter.format("\\draw[color=lightgray] (%s, %s) grid[xstep = %s, ystep=%s] (%s, %s);%n",
            (Math.min(XAxis.getAxis().getRange().getLowerBound(), XAxis.getTwinAxisPosition())-XAxis.getTwinAxisPosition()) * XScale,
            (Math.min(YAxis.getAxis().getRange().getLowerBound(), YAxis.getTwinAxisPosition())-YAxis.getTwinAxisPosition()) * YScale,
            xstepGrid*XScale, ystepGrid*YScale,
            (Math.max(XAxis.getAxis().getRange().getUpperBound(), XAxis.getTwinAxisPosition())-XAxis.getTwinAxisPosition()) * XScale,
            (Math.max(YAxis.getAxis().getRange().getUpperBound(), YAxis.getTwinAxisPosition())-YAxis.getTwinAxisPosition()) * YScale );

      formatter.format("%s", XAxis.toLatex(XScale) );
      formatter.format("%s", YAxis.toLatex(YScale) );

      for(int i = 0; i < datasetList.size(); i++)
         formatter.format("%s", datasetList.get(i).toLatex(XScale, YScale,
                                                            XAxis.getTwinAxisPosition(), YAxis.getTwinAxisPosition(),
                                                            XAxis.getAxis().getLowerBound(), XAxis.getAxis().getUpperBound(),
                                                            YAxis.getAxis().getLowerBound(), YAxis.getAxis().getUpperBound()));

      formatter.format("\\end{tikzpicture}%n");
      formatter.format("\\end{center}%n");
      if (chart.getTitle() != null) {
         formatter.format("\\caption{");
         formatter.format(chart.getTitle().getText());
         formatter.format("}%n\\end{figure}%n");
      }
      if (latexDocFlag)
      formatter.format("\\end{document}%n");
      return formatter.toString();
   }\end{hide}
\end{code}
\begin{tabb}
Same as in \class{XYChart}.
\end{tabb}
\begin{htmlonly}
   \param{width}{Chart's width in centimeters.}
   \param{height}{Chart's height in centimeters.}
\end{htmlonly}
\begin{code}

   public void setLatexDocFlag (boolean flag) \begin{hide} {
      latexDocFlag = flag;
   }\end{hide}
\end{code}
\begin{tabb}
Same as in \class{XYChart}.
\end{tabb}
\begin{code}
\begin{hide}

   protected double computeXScale (double position) {
      double[] bounds = new double[2];
      bounds[0] = XAxis.getAxis().getLowerBound();
      bounds[1] = XAxis.getAxis().getUpperBound();

      if(position<bounds[0])
         bounds[0] = position;
      if(position>bounds[1])
         bounds[1] = position;
      bounds[0] -= position;
      bounds[1] -= position;
      return computeScale(bounds);
   }

   protected double computeYScale (double position) {
      double[] bounds = new double[2];
      bounds[0] = YAxis.getAxis().getLowerBound();
      bounds[1] = YAxis.getAxis().getUpperBound();

      if(position<bounds[0])
         bounds[0] = position;
      if(position>bounds[1])
         bounds[1] = position;
      bounds[0] -= position;
      bounds[1] -= position;
      return computeScale(bounds);
   }


   protected double computeScale (double[] bounds) {
      int tenPowerRatio=0;
      //echelle < 1 si les valeurs sont grandes
      while(bounds[1] > 1000 || bounds[0] < -1000) {
         bounds[1] /= 10;
         bounds[0] /= 10;
         tenPowerRatio++;
      }
      //echelle > 1 si les valeurs sont petites
      while(bounds[1]<100 && bounds[0] > -100) {
         bounds[1] *= 10;
         bounds[0] *= 10;
         tenPowerRatio--;
      }
      return 1/Math.pow(10, tenPowerRatio);
   }

   private static double max (double[] t) {
      double aux = t[0];
      for (int i=1 ; i < t.length ; i++)
         if (t[i] > aux)
            aux = t[i];
      return aux ;
   }

   private static double min (double[] t) {
      double aux = t[0];
      for (int i=1 ; i < t.length ; i++)
         if (t[i] < aux)
            aux = t[i];
      return aux ;
   }
}\end{hide}
\end{code}
