\defclass {ScatterChart}

% Extends \externalclass{umontreal.iro.lecuyer.charts}{XYChart}.
This class provides tools to create and manage scatter plots. Using the
\class{ScatterChart} class is the simplest way to produce scatter plots only.
Each \class{ScatterChart} object is linked with a
\externalclass{umontreal.iro.lecuyer.charts}{XYListSeriesCollection} data set.

\bigskip\hrule
\begin{code}
\begin{hide}
/*
 * Class:        ScatterChart
 * Description:  
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.charts;\begin{hide}

import   org.jfree.chart.axis.NumberAxis;
import   org.jfree.chart.ChartFactory;
import   org.jfree.chart.ChartPanel;
import   org.jfree.chart.plot.XYPlot;
import   org.jfree.chart.plot.PlotOrientation;
import   org.jfree.chart.renderer.xy.XYLineAndShapeRenderer;
import   org.jfree.chart.renderer.xy.XYDotRenderer;
import   org.jfree.data.xy.XYSeriesCollection;
import   java.util.Locale;
import   java.util.Formatter;
import   cern.colt.list.DoubleArrayList;
import   javax.swing.JFrame;
\end{hide}

public class ScatterChart extends XYChart \begin{hide} {

   protected void init (String title, String XLabel, String YLabel) {
      // create the chart...
      chart = ChartFactory.createScatterPlot (
         title,                    // chart title
         XLabel,                   // x axis label
         YLabel,                   // y axis label
         dataset.getSeriesCollection(), // data
         PlotOrientation.VERTICAL,
         true,                     // include legend
         true,                     // tooltips
         false                     // urls
      );

      ((XYPlot)chart.getPlot()).setRenderer(dataset.getRenderer());
      // Initialize axis variables
      initAxis();

      int nb = getSeriesCollection().getSeriesCollection().getSeriesCount();
      for (int i = 0 ; i < nb ; i++) {
         getSeriesCollection().setDashPattern(i, "only marks");
         getSeriesCollection().setMarksType(i, "+");
      }
   }

   protected void initAxis(){
      XAxis = new Axis((NumberAxis)((XYPlot)chart.getPlot()).getDomainAxis(),
                        Axis.ORIENTATION_HORIZONTAL);
      YAxis = new Axis((NumberAxis)((XYPlot)chart.getPlot()).getRangeAxis(),
                        Axis.ORIENTATION_VERTICAL);
      setAutoRange(true, true);
   }
\end{hide}
\end{code}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Constructors}

\begin{code}

   public ScatterChart() \begin{hide} {
      super();
      dataset = new XYListSeriesCollection();
      init (null, null, null);
   }\end{hide}
\end{code}
\begin{tabb}
   Initializes a new \texttt{ScatterChart} instance with an empty data set.
\end{tabb}
\begin{code}

   public ScatterChart (String title, String XLabel, String YLabel,
                        double[][]... data) \begin{hide} {
      super();
      dataset = new XYListSeriesCollection(data);
      init (title, XLabel, YLabel);
   }\end{hide}
\end{code}
\begin{tabb}
Initializes a new \texttt{ScatterChart} instance with data \texttt{data}.
\texttt{title} is a title, \texttt{XLabel} is a short description of the
$x$-axis and \texttt{YLabel} a short description of the $y$-axis.
The input parameter \texttt{data}  represents sets of plotting data.
%
 For example, if one $n$-row matrix \texttt{data1} is given as argument
 \texttt{data}, then the first row \texttt{data1}$[0]$ represents the
 $x$-coordinate vector, and every other row \texttt{data1}$[i],
   i=1,\ldots, n-1$, represents the $y$-coordinates of a set of points.
  Therefore matrix \texttt{data1}  corresponds
   to $n-1$ sets of points, all with the same $x$-coordinates.
%
  However, one may want to plot sets of points with different $x$-coordinates.
  In that case, one should give the points as matrices with two rows.
For examples, if the argument \texttt{data} is made of three 2-row matrices
\texttt{data1}, \texttt{data2} and \texttt{data3}, then they represents
 three different sets of points, \texttt{data*}$[0]$ giving the $x$-coordinates,
 and  \texttt{data*}$[1]$  the $y$-coordinates of the points.
\end{tabb}
\begin{htmlonly}
   \param{title}{chart title.}
   \param{XLabel}{Label on $x$-axis.}
   \param{YLabel}{Label on $y$-axis.}
   \param{data}{series of point sets.}
\end{htmlonly}
\begin{code}

   public ScatterChart (String title, String XLabel, String YLabel,
                        double[][] data, int numPoints) \begin{hide} {
      super();
      dataset = new XYListSeriesCollection(data, numPoints);
      init (title, XLabel, YLabel);
   }\end{hide}
\end{code}
\begin{tabb}
Initializes a new \texttt{ScatterChart} instance with sets of points \texttt{data}.
\texttt{title} is a title, \texttt{XLabel} is a short description of the
$x$-axis, and \texttt{YLabel} a short description of the $y$-axis.
 If \texttt{data} is a $n$-row matrix,
 then the first row \texttt{data}$[0]$ represents the
 $x$-coordinate vector, and every other row \texttt{data}$[i],
   i=1,\ldots, n-1$, represents a $y$-coordinate vector.
  Therefore matrix \texttt{data}$[i][\ ]$, $i=0,\ldots, n-1$,  corresponds
   to $n-1$ sets of points, all with the same $x$-coordinates.
  However, only \emph{the first} \texttt{numPoints} of each set \texttt{data}$[i]$
  (i.e. the first \texttt{numPoints} columns of each row)
  will be plotted.
\end{tabb}
\begin{htmlonly}
   \param{title}{chart title.}
   \param{XLabel}{Label on $x$-axis.}
   \param{YLabel}{Label on $y$-axis.}
   \param{data}{series of point sets.}
   \param{numPoints}{Number of points to plot}
\end{htmlonly}
\begin{code}

   public ScatterChart (String title, String XLabel, String YLabel,
                        double[][] data, int x, int y) \begin{hide} {
      super();
      int len = data[0].length;
      double[][] proj = new double[2][len];
      for (int i = 0; i < len; i++) {
         proj[0][i] = data[x][i];
         proj[1][i] = data[y][i];
      }
      dataset = new XYListSeriesCollection(proj);
      init (title, XLabel, YLabel);
   }\end{hide}
\end{code}
\begin{tabb}
Initializes a new \texttt{ScatterChart} instance using subsets of \texttt{data}.
\texttt{data[x][.]} will form the $x$-coordinates and
\texttt{data[y][.]} will form the $y$-coordinates of the chart.
\texttt{title} sets a title, \texttt{XLabel} is a short description of the
$x$-axis, and \texttt{YLabel} is a short description of the $y$-axis.
Warning: if the new $x$-axis coordinates are not monotone increasing, then
they will automatically be sorted in increasing order so the points will
be reordered, but the original \texttt{data} is not changed.
\end{tabb}
\begin{htmlonly}
   \param{title}{chart title.}
   \param{XLabel}{Label on $x$-axis.}
   \param{YLabel}{Label on $y$-axis.}
   \param{data}{series of point sets.}
   \param{x}{Index of data forming the $x$-coordinates}
   \param{y}{Index of data forming the $y$-coordinates}
\end{htmlonly}
\begin{code}

   public ScatterChart (String title, String XLabel, String YLabel,
                        DoubleArrayList... data) \begin{hide} {
      super();
      dataset = new XYListSeriesCollection(data);
      init (title, XLabel, YLabel);
   }\end{hide}
\end{code}
\begin{tabb}
   Initializes a new \texttt{ScatterChart} instance with data \texttt{data}.
   The input parameter \texttt{data} represents a set of plotting data. A
   \externalclass{cern.colt.list}{DoubleArrayList} from the Colt library is
   used to store the data. The description is similar to the above
   constructor with \texttt{double[]... data}.
% Therefore \texttt{data}$[0]$ will form the first
 %   curve, \texttt{data}$[1]$ forms the second and in general,
%    \texttt{data}$[i], i = 0, \ldots,n,$ forms a point set.
\end{tabb}
\begin{htmlonly}
   \param{title}{chart title.}
   \param{XLabel}{Label on $x$-axis.}
   \param{YLabel}{Label on $y$-axis.}
   \param{data}{series of point sets.}
\end{htmlonly}
\begin{code}

   public ScatterChart (String title, String XLabel, String YLabel,
                        XYSeriesCollection data) \begin{hide} {
      super();
      dataset = new XYListSeriesCollection(data);
      init (title, XLabel, YLabel);
   }\end{hide}
\end{code}
\begin{tabb}
   Initializes a new \texttt{ScatterChart} instance with data \texttt{data}.
   The input parameter \texttt{data} represents a set of plotting data.
   \externalclass{org.jfree.data.xy}{XYSeriesCollection} is a
   \texttt{JFreeChart} container class to store $XY$ plots.
\end{tabb}
\begin{htmlonly}
   \param{title}{chart title.}
   \param{XLabel}{Label on $x$-axis.}
   \param{YLabel}{Label on $y$-axis.}
   \param{data}{series collection.}
\end{htmlonly}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Methods}

\begin{code}

   public int add (double[] x, double[] y, String name, String plotStyle) \begin{hide} {
      int seriesIndex = add(x,y);
      getSeriesCollection().setName(seriesIndex, name);
      getSeriesCollection().setPlotStyle (seriesIndex, plotStyle);
      return seriesIndex;
   }\end{hide}
\end{code}
\begin{tabb}
   Adds a data series into the series collection. Vector \texttt{x} represents
   the $x$-coordinates and vector \texttt{y} represents the $y$-coordinates of
   the series. \texttt{name} and \texttt{plotStyle} are the name and the plot
   style associated to the series.
\end{tabb}
\begin{htmlonly}
   \param{x}{$x_i$ coordinates.}
   \param{y}{$y_i$ coordinates.}
   \param{name}{Name of the series.}
   \param{plotStyle}{Plot style of the series.}
   \return{Integer that represent the new point set's position in the JFreeChart \texttt{XYSeriesCollection} object.}
\end{htmlonly}
\begin{code}

   public int add (double[] x, double[] y) \begin{hide} {
      return add (x, y, x.length);
   }\end{hide}
\end{code}
\begin{tabb}
   Adds a data series into the series collection. Vector \texttt{x} represents
   the $x$-coordinates and vector \texttt{y} represents the $y$-coordinates of
   the series.
\end{tabb}
\begin{htmlonly}
   \param{x}{$x_i$ coordinates.}
   \param{y}{$y_i$ coordinates.}
   \return{Integer that represent the new point set's position in the JFreeChart \texttt{XYSeriesCollection} object.}
\end{htmlonly}
\begin{code}

   public int add (double[] x, double[] y, int numPoints) \begin{hide} {
      int seriesIndex = getSeriesCollection().add(x, y, numPoints);
      initAxis();
      getSeriesCollection().setMarksType(seriesIndex, "+");
      getSeriesCollection().setDashPattern(seriesIndex, "only marks");
      return seriesIndex;
   }\end{hide}
\end{code}
\begin{tabb}
   Adds a data series into the series collection. Vector \texttt{x} represents
   the $x$-coordinates and vector \texttt{y} represents the $y$-coordinates of
   the series. Only \emph{the first} \texttt{numPoints} of \texttt{x} and
   \texttt{y} will be taken into account for the new series.
\end{tabb}
\begin{htmlonly}
   \param{x}{$x_i$ coordinates.}
   \param{y}{$y_i$ coordinates.}
   \param{numPoints}{Number of points to add.}
   \return{Integer that represent the new point set's position in the JFreeChart \texttt{XYSeriesCollection} object.}
\end{htmlonly}
\begin{code}

   public XYListSeriesCollection getSeriesCollection() \begin{hide} {
      return (XYListSeriesCollection)dataset;
   }\end{hide}
\end{code}
\begin{tabb}
   Returns the chart's dataset.
\end{tabb}
\begin{htmlonly}
   \return{the chart's dataset.}
\end{htmlonly}
\begin{code}

   public void setSeriesCollection (XYListSeriesCollection dataset) \begin{hide} {
      this.dataset = dataset;
   }\end{hide}
\end{code}
\begin{tabb}
   Links a new dataset to the current chart.
\end{tabb}
\begin{htmlonly}
   \param{dataset}{new dataset.}
\end{htmlonly}
\begin{code}

   public void setTicksSynchro (int s) \begin{hide} {
      XYSeriesCollection seriesCollection =
          (XYSeriesCollection)this.dataset.getSeriesCollection();
      double[] values = new double[seriesCollection.getItemCount(s)];

      for(int i = 0; i < seriesCollection.getItemCount(s); i++)
         values[i] = seriesCollection.getXValue(s, i);

      XAxis.setLabels(values);
   }\end{hide}
\end{code}
\begin{tabb}
   Synchronizes $X$-axis ticks to the $s$-th series $x$-values.
\end{tabb}
\begin{htmlonly}
   \param{s}{series used to define ticks.}
\end{htmlonly}
\begin{code}

   public JFrame view (int width, int height) \begin{hide} {
      JFrame myFrame;
      if (chart.getTitle () != null)
         myFrame = new JFrame ("ScatterChart from SSJ: " + chart.getTitle ().getText ());
      else
         myFrame = new JFrame ("ScatterChart from SSJ");
      XYPlot plot = chart.getXYPlot ();

/*    // The drawn points are somewhat big, of different shapes, unfilled
      XYLineAndShapeRenderer shape = new XYLineAndShapeRenderer(false, true);
      int nb = getSeriesCollection().getSeriesCollection().getSeriesCount();
      for (int i = 0 ; i < nb ; i++) {
         shape.setSeriesShapesFilled(i, false);
         plot.setRenderer(i, shape);
      }
*/
      // The drawn points are all square, filled
      XYDotRenderer shape = new XYDotRenderer();
      final int dotSize = 3;
      shape.setDotWidth(dotSize);
      shape.setDotHeight(dotSize);
      int nb = getSeriesCollection().getSeriesCollection().getSeriesCount();
      for (int i = 0 ; i < nb ; i++)
         plot.setRenderer(i, shape);

      ChartPanel chartPanel = new ChartPanel (chart);
      chartPanel.setPreferredSize (new java.awt.Dimension(width, height));
      myFrame.setContentPane (chartPanel);
      myFrame.pack();
      myFrame.setDefaultCloseOperation (JFrame.DISPOSE_ON_CLOSE);
      myFrame.setLocationRelativeTo (null);
      myFrame.setVisible (true);
      return myFrame;
   }\end{hide}
\end{code}
\begin{tabb}
   Displays chart on the screen using Swing.
   This method creates an application containing a chart panel displaying
   the chart. The created frame is positioned on-screen, and displayed before
   it is returned. The \texttt{width} and the \texttt{height}
   of the chart are measured in pixels.
\end{tabb}
\begin{htmlonly}
   \param{width}{frame width in pixels.}
   \param{height}{frame height in pixels.}
   \return{frame containing the chart.};
\end{htmlonly}


%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{LaTex-specific method}

\begin{code}

   public String toLatex (double width, double height) \begin{hide} {
      double xunit=0, yunit=0;
      double[] save = new double[4];

      if(dataset.getSeriesCollection().getSeriesCount() == 0)
         throw new IllegalArgumentException("Empty chart");

      //Calcul des parametres d'echelle et de decalage
      double XScale = computeXScale(XAxis.getTwinAxisPosition());
      double YScale = computeYScale(YAxis.getTwinAxisPosition());

         //taille d'une unite en x et en cm dans l'objet "tikzpicture"
      xunit = width / ((Math.max(XAxis.getAxis().getRange().getUpperBound(),
                            XAxis.getTwinAxisPosition()) * XScale)
             - (Math.min(XAxis.getAxis().getRange().getLowerBound(),
                    XAxis.getTwinAxisPosition()) * XScale));
         //taille d'une unite en y et en cm dans l'objet "tikzpicture"
      yunit = height / ((Math.max(YAxis.getAxis().getRange().getUpperBound(),
                         YAxis.getTwinAxisPosition()) * YScale)
           - (Math.min(YAxis.getAxis().getRange().getLowerBound(),
                  YAxis.getTwinAxisPosition()) * YScale));

      Formatter formatter = new Formatter(Locale.US);

      /*Entete du document*/
      if (latexDocFlag) {
         formatter.format("\\documentclass[12pt]{article}%n%n");
         formatter.format("\\usepackage{tikz}%n\\usetikzlibrary{plotmarks}%n\\begin{document}%n%n");
      }
      if(chart.getTitle() != null)
         formatter.format("%% PGF/TikZ picture from SSJ: %s%n", chart.getTitle().getText());
      else
         formatter.format("%% PGF/TikZ picture from SSJ %n");
      formatter.format("%% XScale = %s,  YScale = %s,  XShift = %s,  YShift = %s%n", XScale, YScale, XAxis.getTwinAxisPosition(), YAxis.getTwinAxisPosition());
      formatter.format("%% Therefore, thisFileXValue = (originalSeriesXValue+XShift)*XScale%n");
      formatter.format("%%        and thisFileYValue = (originalSeriesYValue+YShift)*YScale%n%n");
      if (chart.getTitle() != null)
         formatter.format("\\begin{figure}%n");
      formatter.format("\\begin{center}%n");
      formatter.format("\\begin{tikzpicture}[x=%scm, y=%scm]%n", xunit, yunit);
      formatter.format("\\footnotesize%n");
      if(grid)
         formatter.format("\\draw[color=lightgray] (%s, %s) grid[xstep = %s, ystep=%s] (%s, %s);%n",
            (Math.min(XAxis.getAxis().getRange().getLowerBound(),
              XAxis.getTwinAxisPosition())-XAxis.getTwinAxisPosition()) * XScale,
            (Math.min(YAxis.getAxis().getRange().getLowerBound(),
             YAxis.getTwinAxisPosition())-YAxis.getTwinAxisPosition()) * YScale,
            xstepGrid*XScale, ystepGrid*YScale,
            (Math.max(XAxis.getAxis().getRange().getUpperBound(),
              XAxis.getTwinAxisPosition())-XAxis.getTwinAxisPosition()) * XScale,
            (Math.max(YAxis.getAxis().getRange().getUpperBound(),
              YAxis.getTwinAxisPosition())-YAxis.getTwinAxisPosition()) * YScale );
      setTick0Flags();
      formatter.format("%s", XAxis.toLatex(XScale) );
      formatter.format("%s", YAxis.toLatex(YScale) );

      formatter.format("%s", dataset.toLatex(XScale, YScale,
            XAxis.getTwinAxisPosition(), YAxis.getTwinAxisPosition(),
            XAxis.getAxis().getLowerBound(), XAxis.getAxis().getUpperBound(),
            YAxis.getAxis().getLowerBound(), YAxis.getAxis().getUpperBound()));

      formatter.format("\\end{tikzpicture}%n");
      formatter.format("\\end{center}%n");
      if (chart.getTitle() != null) {
         formatter.format("\\caption{");
         formatter.format(chart.getTitle().getText());
         formatter.format("}%n\\end{figure}%n");
      }
      if (latexDocFlag)
         formatter.format("\\end{document}%n");
      return formatter.toString();
   }\end{hide}
\end{code}
\begin{code}
\begin{hide}
}\end{hide}
\end{code}
