\defclass{ListOfTallies}

Represents a list of tally statistical collectors.
Each element of the list is an instance of \externalclass{umontreal.iro.lecuyer.stat}{Tally},
and a vector of observations can be added with
the \method{add}{(double[])} method.
This class defines factory methods to fill a newly-constructed list
with \texttt{Tally} or \texttt{TallyStore} instances.

\bigskip\hrule

\begin{code}
\begin{hide}
/*
 * Class:        ListOfTallies
 * Description:  List of statistical collectors.
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       Éric Buist 
 * @since        2007

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.stat.list;\begin{hide}

import umontreal.iro.lecuyer.util.PrintfFormat;
import cern.colt.list.DoubleArrayList;
import cern.colt.matrix.DoubleMatrix1D;
import cern.colt.matrix.DoubleMatrix2D;

import umontreal.iro.lecuyer.stat.Tally;
import umontreal.iro.lecuyer.stat.TallyStore;
\end{hide}


public class ListOfTallies<E extends Tally> extends ListOfStatProbes<E>\begin{hide} {
\end{hide}
\end{code}
\subsubsection*{Constructors}
\begin{code}

   public ListOfTallies()\begin{hide} {
      super();
   }\end{hide}
\end{code}
\begin{tabb}   Constructs a new empty list of tallies.
\end{tabb}
\begin{code}

   public ListOfTallies (String name)\begin{hide} {
      super (name);
   }\end{hide}
\end{code}
\begin{tabb}   Constructs a new empty list of tallies with name \texttt{name}.
\end{tabb}
\begin{htmlonly}
   \param{name}{the name of the new list.}
\end{htmlonly}
\subsubsection*{Methods}
\begin{code}

   public static ListOfTallies<Tally> createWithTally (int size)\begin{hide} {
      ListOfTallies<Tally> list = new ListOfTallies<Tally>();
      for (int i = 0; i < size; i++)
         list.add (new Tally());
      return list;
   }\end{hide}
\end{code}
\begin{tabb}  This factory method constructs and returns a list of tallies with \texttt{size} instances of
   \class{Tally}.
\end{tabb}
\begin{htmlonly}
   \param{size}{the size of the list.}
   \return{the created list.}
\end{htmlonly}
\begin{code}

   public static ListOfTallies<TallyStore> createWithTallyStore (int size)\begin{hide} {
      ListOfTallies<TallyStore> list = new ListOfTallies<TallyStore>();
      for (int i = 0; i < size; i++)
         list.add (new TallyStore());
      return list;
   }\end{hide}
\end{code}
\begin{tabb}  This factory method constructs and returns a list of tallies with \texttt{size} instances of
   \class{TallyStore}.
\end{tabb}
\begin{htmlonly}
   \param{size}{the size of the list.}
   \return{the created list.}
\end{htmlonly}
\begin{code}

   public void add (double[] x)\begin{hide} {
      int l = size();
      if (x.length != l)
         throw new IllegalArgumentException
            ("Incompatible array length: given " +
            x.length + ", required " + l);
      if (collect)
         for (int i = 0; i < l; i++) {
            double v = x[i];
            Tally ta = get (i);
            if (!Double.isNaN (v) && ta != null)
               ta.add (v);
         }
         notifyListeners (x);
   }\end{hide}
\end{code}
\begin{tabb}   Adds the observation \texttt{x[i]} in
 tally \texttt{i} of this list, for \texttt{i = 0,\ldots, size() - 1}.
 No observation is added if the value is \texttt{Double.NaN},
 or if collecting is turned OFF.
 If broadcasting is ON, the given array is notified
 to all registered observers.
 The given array \texttt{x} not being stored by this object,
 it can be freely used and modified after the call to this method.
\end{tabb}
\begin{htmlonly}
   \param{x}{the array of observations.}
   \exception{NullPointerException}{if \texttt{x} is \texttt{null}.}
   \exception{IllegalArgumentException}{if the length of
    \texttt{x} does not correspond to \texttt{size()}.}
\end{htmlonly}
\begin{code}

   public int numberObs()\begin{hide} {
      if (size() == 0)
         return 0;
      Tally t0 = get (0);
      return t0 == null ? 0 : t0.numberObs();
   }\end{hide}
\end{code}
\begin{tabb}   Assuming that each tally in this list contains
 the same number of observations, returns
 the number of observations in tally~0, or
  0 if this list is empty.
\end{tabb}
\begin{htmlonly}
   \return{the number of observations.}
\end{htmlonly}
\begin{code}

   public boolean areAllNumberObsEqual()\begin{hide} {
      final int l = size();
      int n = numberObs();
      for (int i = 1; i < l; i++) {
         Tally t = get (i);
         if (t.numberObs() != n)
            return false;
      }
      return true;
   }\end{hide}
\end{code}
\begin{tabb}   Tests that every tally in this list contains the
 same number of observations.
 This returns \texttt{true} if and only if all
 tallies have the same number of observations, or if this list is empty.
 If observations are always added using the
 \method{add}{(double[])} method from this class, and not
 \externalmethod{umontreal.iro.lecuyer.stat}{Tally}{add}{(double)} from
   \class{Tally}, this method always  returns \texttt{true}.
\end{tabb}
\begin{htmlonly}
   \return{the success indicator of the test.}
\end{htmlonly}
\begin{code}

   public void average (double[] r)\begin{hide} {
      final int l = size();
      for (int i = 0; i < l; i++) {
          // Manual check to avoid repetitive logs when all tallies
          // have 0 observation.
         Tally ta = get (i);
         double v = ta == null || ta.numberObs() == 0 ? Double.NaN : ta.average();
         r[i] = v;
      }
   }\end{hide}
\end{code}
\begin{tabb}   Computes the average for each tally
 in this list, and stores the averages in the array \texttt{r}.
 If the tally \texttt{i} has no observation,
 the \texttt{Double.NaN} value is stored
 in the array, at index~\texttt{i}.
\end{tabb}
\begin{code}

   public void variance (double[] v)\begin{hide} {
      if (size() != v.length)
         throw new IllegalArgumentException
            ("Invalid length of given array");
      for (int i = 0; i < v.length; i++) {
         Tally tally = get (i);
         if (tally == null || tally.numberObs() < 2)
            v[i] = Double.NaN;
         else
            v[i] = tally.variance();
      }
   }\end{hide}
\end{code}
\begin{tabb}   For each tally in this list, computes
 the sample variance, and stores the variances into the array \texttt{v}.
 If, for some tally~\texttt{i}, there are not enough
 observations for estimating the variance,
 \texttt{Double.NaN} is stored in the array.
\end{tabb}
\begin{htmlonly}
   \param{v}{the array to be filled with sample variances.}
   \exception{NullPointerException}{if \texttt{v} is \texttt{null}.}
   \exception{IllegalArgumentException}{if \texttt{v.length}
    does not correspond to \externalmethod{umontreal.iro.lecuyer.stat.list}{ListOfStatProbes}{size}{()}.}
\end{htmlonly}
\begin{code}

   public void standardDeviation (double[] std)\begin{hide} {
      if (size() != std.length)
         throw new IllegalArgumentException
            ("Invalid length of given array");
      for (int i = 0; i < std.length; i++) {
         Tally tally = get (i);
         if (tally == null || tally.numberObs() < 2)
            std[i] = Double.NaN;
         else
            std[i] = tally.standardDeviation();
      }
   }\end{hide}
\end{code}
\begin{tabb}   For each tally in this list, computes
 the sample standard deviation, and stores the standard deviations
 into the array \texttt{std}.
 This is equivalent to calling \method{variance}{(double[])} and
 performing a square root on every element
 of the filled array.
\end{tabb}
\begin{htmlonly}
   \param{std}{the array to be filled with standard deviations.}
   \exception{NullPointerException}{if \texttt{std} is \texttt{null}.}
   \exception{IllegalArgumentException}{if \texttt{std.length}
    does not correspond to \texttt{size()}.}
\end{htmlonly}
\begin{code}

   public double covariance (int i, int j)\begin{hide} {
      if (i == j)
         return get (i).variance();

      TallyStore tallyi = (TallyStore)get (i);
      TallyStore tallyj = (TallyStore)get (j);
      return tallyi.covariance (tallyj);
   }\end{hide}
\end{code}
\begin{tabb}   Returns the empirical covariance of the observations in tallies
 with indices \texttt{i} and \texttt{j}.  If $x_1,\ldots,x_n$ represent the
 observations in tally \texttt{i} whereas $y_1,\ldots,y_n$ represent the
 observations in tally \texttt{j}, then the covariance is given by
 \[ S_{X, Y} =
 \frac{1}{n-1}\sum_{k=1}^n (x_k - \bar{X}_n)(y_k - \bar{Y}_n)
 = \frac{1}{n-1}\left(\sum_{k=1}^n x_ky_k - \frac{1}{n}
  \sum_{k=1}^n x_k\sum_{r=1}^n y_r\right).\]
   This returns \texttt{Double.NaN}
   if the tallies do not contain the same number of observations, or
 if they contain less than two observations.
 This method throws an exception if the
 underlying tallies are not capable of storing
 observations, i.e. if the tallies are not TallyStores.
 The \class{ListOfTalliesWithCovariance}
 subclass provides an alternative implementation
 of this method which does not require the
 observations to be stored.
\end{tabb}
\begin{htmlonly}
   \param{i}{the index of the first tally.}
   \param{j}{the index of the second tally.}
   \return{the value of the covariance.}
   \exception{ArrayIndexOutOfBoundsException}{if one or both
    indices are out of bounds.}
\end{htmlonly}
\begin{code}

   public double correlation (int i, int j)\begin{hide} {
      if (i == j)
         return 1.0;
      double cov = covariance (i, j);
      Tally tallyi = get (i);
      Tally tallyj = get (j);
      if (tallyi == null || tallyj == null)
         return Double.NaN;
      return cov/Math.sqrt (tallyi.variance()*tallyj.variance());
   }\end{hide}
\end{code}
\begin{tabb}   Returns the empirical correlation between
 the observations in tallies with indices \texttt{i} and \texttt{j}.
 If the tally \texttt{i} contains a sample of the random
 variate $X$ and the tally \texttt{j} contains a sample of $Y$,
 this corresponds to
 \[\mathrm{Cor}(X, Y)=\Cov(X, Y)
 /\sqrt{\Var(X)\Var(Y)}.\]
 
 This method uses \method{covariance}{(int, int)}
 to obtain an estimate of the covariance, and
 \externalmethod{umontreal.iro.lecuyer.stat}{Tally}{variance}{()} in
class \class{Tally} to obtain the sample variances.
\end{tabb}
\begin{htmlonly}
   \param{i}{the index of the first tally.}
   \param{j}{the index of the second tally.}
   \return{the value of the correlation.}
   \exception{ArrayIndexOutOfBoundsException}{if one or both
    indices are out of bounds.}
\end{htmlonly}
\begin{code}

   public void covariance (DoubleMatrix2D c)\begin{hide} {
      int l = size();
      if (c.rows() != l)
         throw new IllegalArgumentException
            ("Invalid number of rows in covariance matrix");
      if (c.columns() != l)
         throw new IllegalArgumentException
            ("Invalid number of columns in covariance matrix");
      for (int i1 = 0; i1 < l; i1++)
         c.setQuick (i1, i1, get (i1).variance());
      for (int i1 = 0; i1 < l - 1; i1++)
         for (int i2 = i1 + 1; i2 < l; i2++) {
            double cov = covariance (i1, i2);
            c.setQuick (i1, i2, cov);
            c.setQuick (i2, i1, cov);
         }
   }\end{hide}
\end{code}
\begin{tabb}   Constructs and returns the sample covariance matrix
 for the tallies in this list.  The given $d\times d$ matrix \texttt{c},
 where $d=$~\texttt{size()},
 is filled with the computed sample covariances.
 Element \texttt{c.get (i, j)} corresponds to
 the result of
 \texttt{covariance (i, j)}.
\end{tabb}
\begin{htmlonly}
   \param{c}{the matrix to be filled with the sample covariances.}
   \exception{NullPointerException}{if \texttt{c} is \texttt{null}.}
   \exception{IllegalArgumentException}{if the number of rows or columns
    in \texttt{c} does not correspond to \texttt{size()}.}
\end{htmlonly}
\begin{code}

   public void correlation (DoubleMatrix2D c)\begin{hide} {
      int l = size();
      if (c.rows() != l)
         throw new IllegalArgumentException
            ("Invalid number of rows in correlation matrix");
      if (c.columns() != l)
         throw new IllegalArgumentException
            ("Invalid number of columns in correlation matrix");
      for (int i1 = 0; i1 < l; i1++)
         c.setQuick (i1, i1, 1.0);
      for (int i1 = 0; i1 < l - 1; i1++)
         for (int i2 = i1 + 1; i2 < l; i2++) {
            double cor = correlation (i1, i2);
            c.setQuick (i1, i2, cor);
            c.setQuick (i2, i1, cor);
         }
   }\end{hide}
\end{code}
\begin{tabb}   Similar to \method{covariance}{(DoubleMatrix2D)} for computing
 the sample correlation matrix.
\end{tabb}
\begin{htmlonly}
   \param{c}{the matrix to be filled with the correlations.}
   \exception{NullPointerException}{if \texttt{c} is \texttt{null}.}
   \exception{IllegalArgumentException}{if the number of rows or columns in \texttt{c}
    does not correspond to \externalmethod{umontreal.iro.lecuyer.stat.list}{ListOfStatProbes}{size}{()}.}
\end{htmlonly}
\begin{code}

   public ListOfTallies<E> clone()\begin{hide} {
      return (ListOfTallies<E>)super.clone();
   }
}\end{hide}
\end{code}
\begin{tabb}   Clones this object.   This makes a shallow copy
  of this list, i.e., this does not clone all the tallies in the list.
  The created clone is modifiable, even if the original list is unmodifiable.
\end{tabb}
