\defmodule {AntitheticStream}

This container class allows the user to force any \class{RandomStream} to 
return antithetic variates.  That is, \method{nextDouble}{} returns 
$1-u$ instead of $u$ and the corresponding change is made in 
\method{nextInt}{}.
Any instance of this class behaves exactly like a \class{RandomStream},
except that it depends on another random number generator stream, called the
\emph{base stream}, to generate its numbers. 
Any call to one of the \texttt{next...}
methods of this class will modify the state of the base stream. 


%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bigskip\hrule

\begin{code}
\begin{hide}
/*
 * Class:        AntitheticStream
 * Description:  container class allows the user to force any RandomStream
                 to return antithetic variates
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.rng; 

public class AntitheticStream implements RandomStream \begin{hide} {

   // The base stream.
   private RandomStream st;

\end{hide}
\end{code}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection* {Constructors}

\begin{code}

   public AntitheticStream (RandomStream stream) \begin{hide} {
      this.st = stream;
   } \end{hide}
\end{code}
\begin{tabb} Constructs a new antithetic stream, using the random numbers
  from the base stream \texttt{stream}.
\end{tabb}

%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection* {Methods}
\begin{code}\begin{hide}
   public void resetStartStream() {
      st.resetStartStream();
   }

   public void resetStartSubstream() {
      st.resetStartSubstream();
   }

   public void resetNextSubstream() {
      st.resetNextSubstream();
   }\end{hide}

   public String toString() \begin{hide} {
      return "Antithetic of " + st.toString();
   }\end{hide}
\end{code}
\begin{tabb} Returns a string starting with \texttt{"Antithetic of "}
  and finishing with the result of the call to the \texttt{toString}
  method of the generator.
\end{tabb}
\begin{code}

   public double nextDouble() \begin{hide} {
      return 1.0 - st.nextDouble();
   }\end{hide}
\end{code}
\begin{tabb} Returns \texttt{1.0 - s.nextDouble()} where \texttt{s} is the 
  base stream.
\end{tabb}
\begin{code}

   public int nextInt (int i, int j) \begin{hide} {
      // pas (j - st.nextInt(0,j-i)), au cas ou le resultat varie.
      return j - i - st.nextInt(i, j);
   }\end{hide}
\end{code}
\begin{tabb} Returns \texttt{j - i - s.nextInt(i, j)} where \texttt{s} is the
  base stream.
\end{tabb}
\begin{code}

   public void nextArrayOfDouble (double[] u, int start, int n) \begin{hide} {
       st.nextArrayOfDouble (u, start, n);
       for (int ii = start; ii < start + n; ii++)
          u[ii] = 1.0 - u[ii];
   }\end{hide}
\end{code} 
\begin{tabb} Calls \texttt{nextArrayOfDouble (u, start, n)} for the base stream,
  then replaces each \texttt{u[i]} by \texttt{1.0 - u[i]}.
\end{tabb}
\begin{htmlonly}
  \param{u}{the array in which the numbers will be stored}
  \param{start}{the first index of \texttt{u} to be used}
  \param{n}{the number of random numbers to put in \texttt{u}}
\end{htmlonly}
\begin{code}

   public void nextArrayOfInt (int i, int j, int[] u, int start, int n) \begin{hide} {
       st.nextArrayOfInt (i, j, u, start, n);
       for (int ii = start; ii < start + n; ii++)
          u[ii] = j - i - u[ii];
   }\end{hide}
\end{code} 
\begin{tabb} Calls \texttt{nextArrayOfInt (i, j, u, start, n)} for the base stream,
  then replaces each \texttt{u[i]} by \texttt{j - i - u[i]}.
\end{tabb}
\begin{htmlonly}
  \param{i}{the smallest possible integer to put in \texttt{u}}
  \param{j}{the largest possible integer to put in \texttt{u}}
  \param{u}{the array in which the numbers will be stored}
  \param{start}{the first index of \texttt{u} to be used}
  \param{n}{the number of random numbers to put in \texttt{u}}
\end{htmlonly}
\begin{code}
\begin{hide}
}
\end{hide}
\end{code}
