\defclass {CloneableRandomStream}

\class{CloneableRandomStream} extends \class{RandomStream} and \class{Cloneable}. 
All classes that implements this interface are able to produce cloned objects.

The cloned object is entirely independent of the older odject.
Moreover the cloned object has all the same properties as the older one. 
All his seeds are duplicated, and therefore both generators will produce the
same random number sequence.

\bigskip\hrule

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{code}
\begin{hide}
/*
 * Class:        CloneableRandomStream
 * Description:  
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.rng;

public interface CloneableRandomStream extends RandomStream, Cloneable \begin{hide} { \end{hide}
\end{code}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection* {Methods}
\begin{code}

   public CloneableRandomStream clone();
\end{code}
 \begin{tabb} Clones the current object and returns its copy.
 \end{tabb}
 \begin{htmlonly}
   \return{A deep copy of the current object}
\end{htmlonly}

\begin{code}\begin{hide} 
}
\end{hide}
\end{code}
