\defclass{RandomStreamInstantiationException}

This exception is thrown when a random stream factory cannot instantiate a stream
on a call to its \externalmethod{umontreal.iro.lecuyer.rng}{RandomStreamFactory}{newInstance}{()} method.

\bigskip\hrule

\begin{code}
\begin{hide}
/*
 * Class:        RandomStreamInstantiationException
 * Description:  thrown when a random stream factory cannot instantiate a stream
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.rng;

public class RandomStreamInstantiationException extends RuntimeException\begin{hide} {
   private RandomStreamFactory factory;
\end{hide}

   public RandomStreamInstantiationException (RandomStreamFactory factory)\begin{hide} {
      super();
      this.factory = factory;
   }\end{hide}
\end{code}
\begin{tabb}   Constructs a new random stream instantiation exception with
 no message, no cause, and thrown by the given \texttt{factory}.
\end{tabb}
\begin{htmlonly}
   \param{factory}{the random stream factory which thrown the exception.}
\end{htmlonly}
\begin{code}

   public RandomStreamInstantiationException (RandomStreamFactory factory,
                                         String message)\begin{hide} {
      super (message);
      this.factory = factory;
   }\end{hide}
\end{code}
\begin{tabb}   Constructs a new random stream instantiation exception with
 the given \texttt{message}, no cause, and concerning \texttt{factory}.
\end{tabb}
\begin{htmlonly}
   \param{factory}{the random stream factory concerned by the exception.}
   \param{message}{the error message describing the exception.}
\end{htmlonly}
\begin{code}

   public RandomStreamInstantiationException (RandomStreamFactory factory,
                                         Throwable cause)\begin{hide} {
      super (cause);
      this.factory = factory;
   }\end{hide}
\end{code}
\begin{tabb}   Constructs a new random stream instantiation exception with
 no message, the given \texttt{cause}, and concerning \texttt{factory}.
\end{tabb}
\begin{htmlonly}
   \param{factory}{the random stream factory concerned by the exception.}
   \param{cause}{the cause of the exception.}
\end{htmlonly}
\begin{code}

   public RandomStreamInstantiationException (RandomStreamFactory factory,
                                         String message, Throwable cause)\begin{hide} {
      super (message, cause);
      this.factory = factory;
   }\end{hide}
\end{code}
\begin{tabb}   Constructs a new random stream instantiation exception with
 the given \texttt{message}, the supplied \texttt{cause}, 
 and concerning \texttt{factory}.
\end{tabb}
\begin{htmlonly}
   \param{factory}{the random stream factory concerned by the exception.}
   \param{message}{the error message describing the exception.}
   \param{cause}{the cause of the exception.}
\end{htmlonly}
\begin{code}

   public RandomStreamFactory getRandomStreamFactory()\begin{hide} {
      return factory;
   }\end{hide}
\end{code}
\begin{tabb}   Returns the random stream factory concerned by this exception.
\end{tabb}
\begin{htmlonly}
   \return{the random stream factory concerned by this exception.}
\end{htmlonly}
\begin{code}

   public String toString()\begin{hide} {
      if (factory == null)
         return super.toString();

      StringBuffer sb = new StringBuffer (getClass().getName());
      sb.append (": For random stream factory ");
      sb.append (factory.toString());
      String msg = getMessage();
      if (msg != null)
         sb.append (", ").append (msg);
      return sb.toString();
   }
}\end{hide}
\end{code}
\begin{tabb}   Returns a short description of the exception.
 If \method{getRandomStreamFactory}{()} returns \texttt{null},
 this calls \texttt{super.toString}. Otherwise, the
 result is the concatenation of:
\end{tabb}
\begin{tabbb}
  a) the name of the actual class of the exception;\\
  b) the string \texttt{": For random stream factory "};\\
  c) the result of \method{getRandomStreamFactory}{()}\texttt{.toString()};\\
  d) if \externalmethod{java.lang}{Throwable}{getMessage}{()} is non-\texttt{null},
     \texttt{", "} followed by the result of 
     \externalmethod{java.lang}{Throwable}{getMessage}{()}.
\end{tabbb}
\begin{htmlonly}
   \return{a string representation of the exception.}
\end{htmlonly}
