\defclass {BinomialGen}

This class implements random variate generators for the 
{\em binomial\/} distribution. It has parameters $n$ and $p$ with
 mass function
\begin{htmlonly}
\eq
     p(x) = \mbox{nCr}(n, x) p^x (1-p)^{n-x} =  
        n!/(x!(n-x)!)\; p^x (1-p)^{n-x} \qquad \mbox {for }
           x=0,1,2,\dots, n
\endeq
where nCr$(n,x)$ is the number of combinations of $x$ objects
among $n$,
\end{htmlonly}
\begin{latexonly}
\eq
     p(x) = {n \choose x} p^x (1-p)^{n-x} =  
        \frac {n!}{x!(n-x)!}\; p^x (1-p)^{n-x} \qquad \mbox {for }
           x=0,1,2,\dots, n                     \label{eq:fmass-binomial}
\endeq
where
\end{latexonly}
$n$ is a positive integer, and $0\le p\le 1$.

% No local copy of the parameters $n$ and $p$ is maintained in this class.
The (non-static) \texttt{nextInt} method simply calls \texttt{inverseF} on the
distribution.
 
\bigskip\hrule

\begin{code}
\begin{hide}
/*
 * Class:        BinomialGen
 * Description:  random variate generators for the binomial distribution
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.randvar;\begin{hide}
import umontreal.iro.lecuyer.probdist.*;
import umontreal.iro.lecuyer.rng.*;\end{hide}

public class BinomialGen extends RandomVariateGenInt \begin{hide} {
   protected int    n = -1;
   protected double p = -1.0;    
    
\end{hide}\end{code}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection* {Constructors}
\begin{code}

   public BinomialGen (RandomStream s, int n, double p)\begin{hide} {
      super (s, new BinomialDist (n, p));
      setParams (n, p);
   }\end{hide}
\end{code}
  \begin{tabb}
  Creates a binomial random variate generator with parameters $n$ and $p$,
  using stream \texttt{s}.
 \end{tabb}
\begin{code}

   public BinomialGen (RandomStream s, BinomialDist dist) \begin{hide} {
      super (s, dist);
      if (dist != null)
         setParams (dist.getN(), dist.getP());
   }\end{hide}
\end{code}
  \begin{tabb} Creates a random variate generator for the {\em binomial\/} 
    distribution \texttt{dist} and the random stream \texttt{s}. 
 \end{tabb}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection* {Methods}
\begin{code}

   public static int nextInt (RandomStream s, int n, double p)\begin{hide} {
      return BinomialDist.inverseF (n, p, s.nextDouble());
   }\end{hide}
\end{code}
\begin{tabb}
   Generates a new integer from the {\em binomial\/} distribution with
  parameters
   $n = $~\texttt{n} and $p = $~\texttt{p}, using the given stream \texttt{s}.
\end{tabb}
\begin{code}

   public int getN()\begin{hide} {
      return n;
   }
   \end{hide}
\end{code}
\begin{tabb} Returns the parameter $n$ of this object.
\end{tabb}
\begin{code}

   public double getP()\begin{hide} {
      return p;
   }
   \end{hide}
\end{code}
\begin{tabb} Returns the parameter $p$ of this object.
\end{tabb}
\begin{hide}\begin{code}

   protected void setParams (int n, double p) {
      if (p < 0.0 || p > 1.0)
         throw new IllegalArgumentException ("p not in range [0, 1]");
      if (n <= 0)
         throw new IllegalArgumentException ("n <= 0");
      this.p = p;
      this.n = n;
   }
\end{code}
\begin{tabb} Sets the parameter $n$ and $p$ of this object.
\end{tabb}
\begin{code}
}
\end{code}
\end{hide}
