 \defclass {InverseFromDensityGen}

Implements a method for generating random variates by numerical inversion of
an \emph{arbitrary continuous} distribution when only the probability density
is known \cite{rDER10a}.  The cumulative probabilities (cdf) are pre-computed by
 numerical quadrature  of the
density using Gauss-Lobatto integration over suitably small intervals to
satisfy the required precision, and these values are kept in tables. Then the
 algorithm uses polynomial interpolation  over the tabulated values to get
  the inverse cdf. The user can select the
  desired precision and the degree of the interpolating polynomials.

The algorithm may fail for some distributions for which the density
 becomes infinite at a point (for ex. the Gamma and the Beta distributions
 with $\alpha < 1$)  if one requires too high a precision
(a  too small \texttt{eps}, for ex. $\epsilon \sim 10^{-15}$).
However, it should work also for continuous densities with finite discontinuities.


While the setup time is relatively slow, the generation of random variables
is extremely fast and practically independent of the required precision
and of the specific distribution. The following table shows the time needed
(in seconds) to generate $10^8$ random numbers using inversion
from a given class, then the numerical inversion with Gauss-Lobatto integration
implemented here, and finally the speed ratios between the two methods.
The speed ratio is  the speed of the latter over the former.
Thus for the beta distribution with parameters (5, 500), generating random
variables with the Gauss-Lobatto integration implemented in this class is
more than 1700
times faster than using inversion from the \texttt{BetaDist} class.
These tests were made on a machine with processor AMD Athlon 4000, running
Red Hat Linux, with clock speed at 2403 MHz.
% \hrichard {The set-up time is not  included in these ratios.}

\begin{center}
\begin{tabular}{|l|c|c|c|}
\hline
 Distribution  & Inversion  & Gauss-Lobatto &  speed ratio  \\
\hline
\texttt{NormalDist(10.5, 5)} & \phantom{1521}9.19 & 8.89 &  \phantom{16222}1.03  \\
\texttt{ExponentialDist(5)}  & \phantom{152}17.72 &8.82 & \phantom{1622}2.0 \\
\texttt{CauchyDist(10.5, 5)} & \phantom{152}18.30 & 8.81 & \phantom{1622}2.1     \\
\texttt{BetaSymmetricalDist(10.5)}   & \phantom{15}242.80 & 8.85 &  \phantom{162}27.4  \\
\texttt{GammaDist(55)} &  \phantom{15}899.50 &8.89 & \phantom{1}101  \\
\texttt{ChiSquareNoncentralDist(10.5, 5)} &\phantom{1}5326.90 &8.85 &  \phantom{1}602  \\
\texttt{BetaDist(5, 500)} & 15469.10 &8.86 &  1746  \\
\hline
\end{tabular}
\end{center}

The following table gives the time (in sec.) needed to
create an object (setup time) and to generate one random variable for
this class compared to the same for the inversion method specific to each class,
and the ratios of the times (init + one random variable) of the two methods.
For inversion, we initialized $10^8$ times; for this class, we
initialized $10^4$ times.

\begin{center}
\begin{tabular}{|l|c|c|c|}
\hline
 Distribution  &Inversion & Gauss-Lobatto &  time ratio  \\
   & $10^8$ init   &  $10^4$ init & for 1 init \\
\hline
\texttt{NormalDist(10.5, 5)} &\phantom{00}5.30 &\phantom{0}38.29 &  26426  \\
\texttt{ExponentialDist(5)}   &\phantom{00}3.98 &\phantom{0}27.05 & 12466 \\
\texttt{CauchyDist(10.5, 5)}  &\phantom{00}5.05 &\phantom{0}58.39 & 25007   \\
\texttt{BetaSymmetricalDist(10.5)}  &\phantom{0}90.66 &  \phantom{0}68.33 & \phantom{0}2049 \\
\texttt{GammaDist(55)}       &\phantom{0}13.15 & \phantom{0}58.34  & \phantom{00}639  \\
\texttt{ChiSquareNoncentralDist(10.5, 5)} &190.48 & 248.98& \phantom{00}451  \\
\texttt{BetaDist(5, 500)}    &\phantom{0}63.60 & 116.57  & \phantom{000}75  \\
\hline
\end{tabular}
\end{center}

If only a few random variables are needed, then using this class
is not efficient because of the slow set-up. But if one wants to generate
large samples from the same distribution with fixed parameters, then
this class will be very efficient. The following table gives the number of
random variables generated beyond which, using this class will be
worthwhile.

\begin{center}
\begin{tabular}{|l|c|}
\hline
 Distribution  &  number of generated variables  \\
\hline
\texttt{NormalDist(10.5, 5)} &  41665  \\
\texttt{ExponentialDist(5)}   & 15266 \\
\texttt{CauchyDist(10.5, 5)}  & 31907   \\
\texttt{BetaSymmetricalDist(10.5)}    & 2814 \\
\texttt{GammaDist(55)}         &  649  \\
\texttt{ChiSquareNoncentralDist(10.5, 5)} &  467 \\
\texttt{BetaDist(5, 500)}      & 75  \\
\hline
\end{tabular}
\end{center}

Thus, for example, if one needs to generate less than 15266 exponential
random variables, then using the \texttt{InverseFromDensityGen} class
is not wortwhile: it will be faster to use inversion from the
\texttt{ExponentialGen} class.


\bigskip\hrule

\begin{code}
\begin{hide}
/*
 * Class:        InverseFromDensityGen
 * Description:  generator of random variates by numerical inversion of
                 an arbitrary continuous distribution when only the 
                 probability density is known
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       Richard Simard
 * @since        June 2008

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.randvar;
   import umontreal.iro.lecuyer.functions.MathFunction;
   import umontreal.iro.lecuyer.rng.RandomStream;
   import umontreal.iro.lecuyer.probdist.ContinuousDistribution;\begin{hide}
   import umontreal.iro.lecuyer.probdist.InverseDistFromDensity;
\end{hide}


public class InverseFromDensityGen extends RandomVariateGen \begin{hide} {
\end{hide}\end{code}

\subsubsection* {Constructors}
\begin{code}

   public InverseFromDensityGen (RandomStream stream,
                                 ContinuousDistribution dis,
                                 double xc, double eps, int order) \begin{hide} {
      super (stream, null);
      dist = new InverseDistFromDensity (dis, xc, eps, order);
   }\end{hide}
\end{code}
\begin{tabb} Creates a new generator for the \emph{continuous} distribution
\texttt{dis}, using stream \texttt{stream}. \texttt{dis} must have a well-defined
density method; its other methods are unused. For a non-standard distribution
\texttt{dis}, the user may wish to set the left and the right boundaries
between which the density is non-zero by calling methods
\externalmethod{umontreal.iro.lecuyer.probdist}{ContinuousDistribution}{setXinf}{}
and
\externalmethod{umontreal.iro.lecuyer.probdist}{ContinuousDistribution}{setXsup}{}
of \texttt{dis}, for better efficiency.
Argument \texttt{xc} can be the mean,
the mode or any other $x$ for which the density is relatively large.
The $u$-resolution \texttt{eps} is the desired absolute error in the CDF,
and \texttt{order} is the degree of the
Newton interpolating polynomial over each interval.
An \texttt{order} of 3 or 5, and an \texttt{eps} of $10^{-6}$ to $10^{-12}$
are usually good choices.
 Restrictions: $3 \le \texttt{order} \le 12$.
\end{tabb}
\begin{code}

   public InverseFromDensityGen (RandomStream stream, MathFunction dens,
                                 double xc, double eps, int order,
                                 double xleft, double xright) \begin{hide} {
      super (stream, null);
      dist = new InverseDistFromDensity (dens, xc, eps, order, xleft, xright);
   } \end{hide}
\end{code}
\begin{tabb} Creates a new generator from the \emph{continuous} probability density
\texttt{dens}. The left and the right boundaries of the density are
\texttt{xleft} and \texttt{xright} (the density is 0 outside the
interval \texttt{[xleft, xright]}).
See the description of the other constructor.
\end{tabb}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\subsubsection* {Methods}
\begin{code}

   public double nextDouble() \begin{hide} {
      return dist.inverseF (stream.nextDouble());
   }\end{hide}
\end{code}
 \begin{tabb} Generates a new random variate.
 \end{tabb}
\begin{code}

   public double getXc()\begin{hide} {
      return ((InverseDistFromDensity)dist).getXc();
   }\end{hide}
\end{code}
\begin{tabb}
   Returns the \texttt{xc} given in the constructor.
\end{tabb}
\begin{code}

   public double getEpsilon()\begin{hide} {
      return ((InverseDistFromDensity)dist).getEpsilon();
   }
\end{hide}
\end{code}
\begin{tabb}
   Returns the $u$-resolution \texttt{eps}.
\end{tabb}
\begin{code}

   public int getOrder()\begin{hide} {
      return ((InverseDistFromDensity)dist).getOrder();
   }
\end{hide}
\end{code}
\begin{tabb}
   Returns the order of the interpolating polynomial.
\end{tabb}
\begin{hide}
\begin{code}
}\end{code}
\end{hide}
