\defclass {NormalGen}

This class implements methods for generating random variates from the
{\em normal\/} distribution $N(\mu, \sigma)$.
It has mean $\mu$ and variance $\sigma^2$, where $\sigma>0$.
 Its density function is
\begin{htmlonly}
\eq
     f(x) = 1/\sqrt{2\pi}\sigma e^{(x-\mu)^2/(2\sigma^2)}
\endeq
\end{htmlonly}
\begin{latexonly}
\eq
     f(x) = \frac{1}{\sqrt{2\pi}\sigma} e^{(x-\mu)^2/(2\sigma^2)}
                        \eqlabel{eq:fnormal}
\endeq
\end{latexonly}
%
The \texttt{nextDouble} method simply calls \texttt{inverseF} on the
distribution.

The following table gives the CPU time needed to generate $10^8$ standard
normal random variates using the different implementations available in SSJ.
The first time is for a generator object (non-static method), and
the second time is for the static method
where no object is created.
These tests were made on a machine with processor AMD Athlon 4000, running
Red Hat Linux, with clock speed at 2403 MHz. The static method
 \texttt{nextDouble()} for \texttt{NormalBoxMullerGen} and
  \texttt{NormalPolarGen} uses only one number out of two that are
  generated; thus they are twice slower than the non-static method.

\begin{center}
\begin{tabular}{|l|c|c|}
\hline
 Generator  &  time in seconds  &  time in seconds  \\
   &  (object)  &  (static)  \\
\hline
\texttt{NormalGen}     &  7.67  & 7.72 \\
\texttt{NormalACRGen}   &    4.71  & 4.76 \\
\texttt{NormalBoxMullerGen}    &  16.07 & 31.45 \\
\texttt{NormalPolarGen}     &  7.31  & 13.74 \\
\texttt{NormalKindermannRamageGen}     & 5.38 & 5.34 \\
\hline
\end{tabular}
\end{center}

\bigskip\hrule

\begin{code}
\begin{hide}
/*
 * Class:        NormalGen
 * Description:  random variates generator from the normal distribution
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.randvar;\begin{hide}
import umontreal.iro.lecuyer.rng.*;
import umontreal.iro.lecuyer.probdist.*;
\end{hide}

public class NormalGen extends RandomVariateGen \begin{hide} {
   protected double mu;
   protected double sigma = -1.0;
\end{hide}\end{code}

\subsubsection* {Constructors}

\begin{code}

   public NormalGen (RandomStream s, double mu, double sigma) \begin{hide} {
      super (s, new NormalDist(mu, sigma));
      setParams (mu, sigma);
   }\end{hide}
\end{code}
\begin{tabb}  Creates a normal random variate generator with mean \texttt{mu}
  and standard deviation \texttt{sigma}, using stream \texttt{s}.
\end{tabb}
\begin{code}

   public NormalGen (RandomStream s) \begin{hide} {
      this (s, 0.0, 1.0);
   }\end{hide}
\end{code}
\begin{tabb}  Creates a standard normal random variate generator with mean
  \texttt{0} and standard deviation \texttt{1}, using stream \texttt{s}.
\end{tabb}
\begin{code}

   public NormalGen (RandomStream s, NormalDist dist) \begin{hide} {
      super (s, dist);
      if (dist != null)
         setParams (dist.getMu(), dist.getSigma());
   }\end{hide}
\end{code}
\begin{tabb} Creates a random variate generator for the normal distribution
  \texttt{dist} and stream \texttt{s}.
\end{tabb}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\subsubsection* {Methods}
\begin{code}

   public static double nextDouble (RandomStream s, double mu, double sigma) \begin{hide} {
      return NormalDist.inverseF (mu, sigma, s.nextDouble());
   }\end{hide}
\end{code}
 \begin{tabb}  Generates a variate from the normal distribution with
   parameters $\mu = $~\texttt{mu} and $\sigma = $~\texttt{sigma}, using
   stream \texttt{s}.
 \end{tabb}
\begin{code}

   public double getMu()\begin{hide} {
      return mu;
   }\end{hide}
\end{code}
  \begin{tabb}  Returns the parameter $\mu$ of this object.
  \end{tabb}
\begin{code}

   public double getSigma()\begin{hide} {
      return sigma;
   }\end{hide}
\end{code}
  \begin{tabb}  Returns the parameter $\sigma$ of this object.
  \end{tabb}
\begin{hide}\begin{code}

   protected void setParams (double mu, double sigma)\begin{hide} {
      if (sigma <= 0)
         throw new IllegalArgumentException ("sigma <= 0");
      this.mu = mu;
      this.sigma = sigma;
   }\end{hide}
\end{code}
  \begin{tabb}  Sets the parameters $\mu$ and $\sigma$ of this object.
  \end{tabb}
\begin{code}
}
\end{code}
\end{hide}
