\defclass{NormalInverseGaussianGen}

\newcommand{\nig}{\emph{normal inverse gaussian}}

This class implements random variate generators for 
the \nig{} ($\mathcal{NIG}$) distribution. See the definition of
\externalclass{umontreal.iro.lecuyer.probdist}{NormalInverseGaussianDist}
\begin{latexonly}%
in package \texttt{probdist}.
\end{latexonly}


\bigskip\hrule\bigskip

\begin{code}
\begin{hide}
/*
 * Class:        NormalInverseGaussianGen
 * Description:  random variate generators for the normal inverse gaussian distribution
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       Richard Simard
 * @since        June 2008

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.randvar;\begin{hide}
import umontreal.iro.lecuyer.rng.*;
import umontreal.iro.lecuyer.probdist.*;
\end{hide}

public class NormalInverseGaussianGen extends RandomVariateGen \begin{hide} {
   protected double mu;
   protected double delta = -1.0;
   protected double alpha = -1.0;
   protected double beta = -2.0;
   protected double gamma = -1.0;
\end{hide}
\end{code}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection* {Constructors}
\begin{code}

   public NormalInverseGaussianGen (RandomStream s, double alpha,
                                    double beta, double mu, double delta) \begin{hide} {
      super (s, new NormalInverseGaussianDist(alpha, beta, mu, delta));
      setParams (alpha, beta, mu, delta);
   }\end{hide}
\end{code} 
\begin{tabb} Creates an \nig{} random variate generator with parameters
 $\alpha$ = \texttt{alpha}, $\beta$ = \texttt{beta},  $\mu$ = \texttt{mu}
 and  $\delta$ = \texttt{delta}, using stream \texttt{s}. 
\end{tabb}
\begin{code}

   public NormalInverseGaussianGen (RandomStream s,
                                    NormalInverseGaussianDist dist) \begin{hide} {
      super (s, dist);
      if (dist != null)
         setParams (dist.getAlpha(), dist.getBeta(), dist.getMu(),
                    dist.getDelta());
   }\end{hide}
\end{code}
  \begin{tabb} Creates a new generator for the distribution \texttt{dist},
     using stream \texttt{s}.
  \end{tabb}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\subsubsection* {Methods}
\begin{code}

   public static double nextDouble (RandomStream s, double alpha,
                                    double beta, double mu, double delta)\begin{hide} {
      return NormalInverseGaussianDist.inverseF (alpha, beta, mu, delta,
                                                 s.nextDouble());
   }\end{hide}
\end{code}
\begin{tabb}
 NOT IMPLEMENTED. Use the daughter classes.
% Generates a variate from the inverse gaussian distribution
%   with location parameter $\mu > 0$ and scale parameter $\lambda > 0$.
\end{tabb}
\begin{code}

   public double getAlpha()\begin{hide} {
      return alpha;
   }\end{hide}
\end{code}
 \begin{tabb} Returns the parameter $\alpha$ of this object.
 \end{tabb}
\begin{code}

   public double getBeta()\begin{hide} {
      return beta;
   }\end{hide}
\end{code}
 \begin{tabb} Returns the parameter $\beta$ of this object.
 \end{tabb}
\begin{code}

   public double getMu()\begin{hide} {
      return mu;
   }\end{hide}
\end{code}
 \begin{tabb} Returns the parameter $\mu$ of this object.
 \end{tabb}
\begin{code}

   public double getDelta()\begin{hide} {
      return delta;
   }\end{hide}
\end{code}
 \begin{tabb} Returns the parameter $\delta$ of this object.
 \end{tabb}
\begin{code}

   public void setParams (double alpha, double beta, double mu,
                          double delta)\begin{hide} {
      if (delta <= 0.0)
         throw new IllegalArgumentException ("delta <= 0");
      if (alpha <= 0.0)
         throw new IllegalArgumentException ("alpha <= 0");
      if (Math.abs(beta) >= alpha)
         throw new IllegalArgumentException ("|beta| >= alpha");

      gamma = Math.sqrt(alpha*alpha - beta*beta);

      this.mu = mu;
      this.delta = delta;
      this.beta = beta;
      this.alpha = alpha;
   }\end{hide}
\end{code}
\begin{tabb}
   Sets the parameters  $\alpha$,  $\beta$, $\mu$ and $\delta$ of this object.
\end{tabb}
\begin{code}\begin{hide} 
}
\end{hide}
\end{code}
