\defclass {NormalInverseFromDensityGen}

This class implements \emph{normal} random variate generators
 using numerical inversion of the \emph{normal} density
 as described in \cite{rDER10a}.  It makes use of the class
\externalclass{umontreal.iro.lecuyer.probdist}{InverseDistFromDensity}.
A set of tables are precomputed to speed up the generation of normal random 
variables by numerical inversion. This will be useful if one 
wants to generate a large number of random variables.


\bigskip\hrule

\begin{code}
\begin{hide}
/*
 * Class:        NormalInverseFromDensityGen
 * Description:  random variate generators using numerical inversion of
                 the normal density
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       Richard Simard
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.randvar;\begin{hide}
import umontreal.iro.lecuyer.rng.*;
import umontreal.iro.lecuyer.probdist.*;
\end{hide}

public class NormalInverseFromDensityGen extends NormalGen \begin{hide} {

\end{hide}\end{code}

\subsubsection* {Constructors}
\begin{code}

   public NormalInverseFromDensityGen (RandomStream stream, double mu, 
                                       double sigma, double ueps, int order) \begin{hide} {
      // dist is the normal distribution
      super (stream, mu, sigma);
      double xc = mu;

      // member (NormalDist) dist is replaced by 
      // (InverseDistFromDensity) dist
      dist = new InverseDistFromDensity ((ContinuousDistribution) dist,
                                         xc, ueps, order);
    }\end{hide}
\end{code} 
\begin{tabb} Creates a normal random variate generator with parameters 
$\mu=$ \texttt{mu} and $\sigma $ = \texttt{sigma}, using stream \texttt{stream}.
It uses numerical inversion with precomputed tables.
The $u$-resolution \texttt{ueps} is the desired absolute error in the 
\texttt{cdf}, and \texttt{order} is the degree of the Newton interpolating 
polynomial over each interval.
\end{tabb}
\begin{code}

   public NormalInverseFromDensityGen (RandomStream stream, NormalDist dist,
                                       double ueps, int order) \begin{hide} {
      super (stream, dist);
      double xc = mu;

      // member (NormalDist) dist is replaced by 
      // (InverseDistFromDensity) dist
      this.dist = new InverseDistFromDensity (dist, xc, ueps, order);
   } \end{hide}
\end{code}
 \begin{tabb}  Similar to the first constructor, with the normal 
   distribution \texttt{dist}.  
 \end{tabb}
\begin{code}

   public NormalInverseFromDensityGen (RandomStream stream, 
                                       InverseDistFromDensity dist) \begin{hide} {
      super (stream, null);
      mu = dist.getXc();
      this.dist = dist;
   } \end{hide}
\end{code}
 \begin{tabb}  Creates a new normal generator using the \emph{normal}
   distribution \texttt{dist} and stream \texttt{stream}. \texttt{dist}
   may be obtained by calling method \method{getDistribution}{()},
   after using one of the other constructors to create the 
   precomputed tables. This is useful when one needs many   generators
 using the same normal distribution.
 Precomputing tables for numerical inversion is
 costly; thus using only one set of tables for many generators 
is more efficient. The first \class{NormalInverseFromDensityGen} generator 
 using the other constructors creates the precomputed tables.
Then all other streams use this constructor with the same set of tables.
\end{tabb}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\subsubsection* {Methods}
\begin{code}

   public double getUepsilon()\begin{hide} {
      return ((InverseDistFromDensity)dist).getEpsilon();
   }
\end{hide}
\end{code}
\begin{tabb}
   Returns the $u$-resolution \texttt{ueps}.
\end{tabb}
\begin{code}

   public int getOrder()\begin{hide} {
      return ((InverseDistFromDensity)dist).getOrder();
   }
\end{hide}
\end{code}
\begin{tabb}
   Returns the order of the interpolating polynomial.
\end{tabb}

\begin{hide}
\begin{code}
}\end{code}
\end{hide}
