\defclass {TriangularGen}

This class implements random variate generators for the 
{\em triangular\/} distribution. Its  density is
\begin{htmlonly}
\[\begin{array}{rll}
f(x) =& 2(x-a)/[(b-a)(m-a)]&\qquad\mbox{ for }a\le x\le m, \\
f(x) =& 2(b-x)/[(b-a)(b-m)]&\qquad\mbox{ for }m\le x\le b, \\
f(x) =& 0&\qquad\mbox{ elsewhere,}
\end{array}\]
\end{htmlonly}
\begin{latexonly}
\eq
    f(x) = \left \{\begin{array}{ll}
           \frac {2(x-a)}{(b-a)(m-a)}  & \mbox {for } a\le x\le m, \\ [6pt]
           \frac {2(b-x)}{(b-a)(b-m)} & \mbox { for } m\le x\le b, \\ [6pt]
            0 & \mbox { elsewhere, }
          \end{array}\right.         \eqlabel{eq:ftrian}
\endeq
\end{latexonly}
where $a\le m\le b$ (see, e.g., \cite{sLAW00a}).

The (non-static) \texttt{nextDouble} method simply calls \texttt{inverseF} on the
distribution.

\bigskip\hrule


\begin{code}
\begin{hide}
/*
 * Class:        TriangularGen
 * Description:  random variate generators for the triangular distribution
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.randvar;\begin{hide}
import umontreal.iro.lecuyer.rng.*;
import umontreal.iro.lecuyer.probdist.*;
\end{hide}

public class TriangularGen extends RandomVariateGen \begin{hide} {
   private double a;
   private double b;
   private double m;
\end{hide}\end{code}

\subsubsection* {Constructors}
\begin{code}

   public TriangularGen (RandomStream s, double a, double b, double m) \begin{hide} {
      super (s, new TriangularDist(a, b, m));
      setParams (a, b, m);
   }\end{hide}
\end{code} 
\begin{tabb}  Creates a triangular random variate generator over the interval
(\texttt{a}, \texttt{b}), with parameter \texttt{m}, using stream \texttt{s}. 
\end{tabb}
\begin{code}

   public TriangularGen (RandomStream s, double m) \begin{hide} {
      this (s, 0.0, 1.0, m);
   }\end{hide}
\end{code} 
\begin{tabb}  Creates a triangular random variate generator over the interval
  $(0, 1)$, with parameter \texttt{m}, using stream \texttt{s}.
\end{tabb}
\begin{code}
 
   public TriangularGen (RandomStream s, TriangularDist dist) \begin{hide} {
      super (s, dist);
      if (dist != null)
         setParams (dist.getA(), dist.getB(), dist.getM());
   }\end{hide}
\end{code}
\begin{tabb}  Creates a new generator for the triangular distribution
 \texttt{dist} and stream \texttt{s}.   
\end{tabb}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\subsubsection* {Methods}
\begin{code}
   
   public static double nextDouble (RandomStream s, double a, 
                                    double b, double m) \begin{hide} {
       // the code is taken and adapted from unuran
       // file /distributions/c_triangular_gen.c
       return TriangularDist.inverseF (a, b, m, s.nextDouble());
   }\end{hide}
\end{code}
\begin{tabb}  
  Generates a new variate from the triangular distribution with parameters
  $a = $~\texttt{a}, $b = $~\texttt{b} and $m = $~\texttt{m} and stream \texttt{s},
   using inversion.
\end{tabb}
\begin{code}      

   public double getA()\begin{hide} {
      return a;
   }\end{hide}
\end{code}
  \begin{tabb}
  Returns the value of $a$ for this object.
 \end{tabb}
\begin{code}      

   public double getB()\begin{hide} {
      return b;
   }\end{hide}
\end{code}
  \begin{tabb}
  Returns the value of $b$ for this object.
 \end{tabb}
\begin{code}      

   public double getM()\begin{hide} {
      return m;
   }\end{hide}
\end{code}
  \begin{tabb}
  Returns the value of $m$ for this object.
 \end{tabb}
\begin{hide}\begin{code}      

   private void setParams (double a, double b, double m) {
      if ((a == 0.0 && b == 1.0) && (m < 0 || m > 1))
         throw new IllegalArgumentException ("m is not in [0,1]");
      else if (a >= b)
         throw new IllegalArgumentException ("a >= b");
      else if (m < a || m > b) 
         throw new IllegalArgumentException ("m is not in [a,b]");
      this.a = a;
      this.b = b;
      this.m = m;
   }
\end{code}
  \begin{tabb}
  Sets the value of the parameters $a$, $b$ and $m$ for this object.
 \end{tabb}
\begin{code}      
}
\end{code}
\end{hide}
