\defclass{PiecewiseConstantFunction}


Represents a piecewise-constant function.

\bigskip\hrule

\begin{code}
\begin{hide}
/*
 * Class:        PiecewiseConstantFunction
 * Description:  
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       Éric Buist
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.functions;\begin{hide}

import java.util.Arrays;
\end{hide}

public class PiecewiseConstantFunction implements MathFunction\begin{hide} {
   private double[] x;
   private double[] y;
\end{hide}

   public PiecewiseConstantFunction (double[] x, double[] y)\begin{hide} {
      if (x.length != y.length)
         throw new IllegalArgumentException();
      this.x = x.clone ();
      this.y = y.clone ();
   }\end{hide}
\end{code}
\begin{tabb}   Constructs a new piecewise-constant function
 with $X$ and $Y$ coordinates given
 by \texttt{x} and \texttt{y}.
\end{tabb}
\begin{htmlonly}
   \param{x}{the $X$ coordinates.}
   \param{y}{the $Y$ coordinates.}
\end{htmlonly}
\begin{code}

   public double[] getX()\begin{hide} {
      return x.clone ();
   }\end{hide}
\end{code}
\begin{tabb}   Returns the $X$ coordinates of the function.
\end{tabb}
\begin{htmlonly}
   \return{the $X$ coordinates of the function.}
\end{htmlonly}
\begin{code}

   public double[] getY()\begin{hide} {
      return y.clone ();
   }\end{hide}
\end{code}
\begin{tabb}   Returns the $Y$ coordinates of
 the function.
\end{tabb}
\begin{htmlonly}
   \return{the $Y$ coordinates of the function.}
\end{htmlonly}
\begin{code}\begin{hide}

   public double evaluate (double x) {
      final int idx = Arrays.binarySearch (this.x, x);
      if (idx >= 0)
         return y[idx];
      final int insertionPoint = -(idx + 1);
      return y[insertionPoint];
   }
}\end{hide}
\end{code}
