\latex{\section*{Overview of package \texttt{stat}}\addcontentsline{toc}{section}{Overview of package \texttt{stat}}}

This package provides elementary tools for collecting statistics and computing
confidence intervals.  The base class 
\externalclass{umontreal.iro.lecuyer.stat}{StatProbe} 
implements common methods needed by all probes.  
Its subclass \externalclass{umontreal.iro.lecuyer.stat}{Tally}
collects data as a sequence of observations $X_1,X_2,\dots$, and computes
sample averages, sample standard deviations, and confidence intervals
based on the normality assumption. \richard{Toujours bas\'e sur la normale?}
The class \externalclass{umontreal.iro.lecuyer.stat}{TallyStore} is 
similar, but it also stores the individual observations in a list 
implemented as a \externalclass{cern.colt.list}{DoubleArrayList}%
\begin{latexonly}, a class imported from the COLT library\end{latexonly}.
This permits one to compute more quantities and to use the methods 
provided by COLT for computing descriptive statistics.

The class \externalclass{umontreal.iro.lecuyer.simevents}{Accumulate}%
\begin{latexonly}, in package \texttt{simevents},\end{latexonly}
computes integrals and averages with respect to time.
This class is in package \texttt{simevents} because its operation depends
on the simulation clock.

\iffalse  %%%%%%%%
The class
\externalclass{umontreal.iro.lecuyer.stat}{FunctionOfMultipleMeansTally}
collects data as a sequence of $d$-dimensional
vectors $\boldX_1, \boldX_2, \ldots$,
and computes a function of sample averages with a sample variance, and
a confidence interval on the corresponding function of means using the
delta theorem.
The function of averages to consider, as well as its gradient, need to
be computed by an implementation of
\externalclass{umontreal.iro.lecuyer.util}{MultivariateFunction},
for example,
\externalclass{umontreal.iro.lecuyer.util}{RatioFunction} for
a ratio of two quantities.
\fi  %%%%%%%%%

All classes that represent statistical probes
support the \emph{observer} design pattern, well-known
in software engineering \cite{iGAM98a}.
%%  Gamma et al., Design Patterns---Elements of Reusable 
%%  Object-Oriented Software, Addison-Wesley, second ed., 1998.
This pattern facilitates the separation of data generation (by the 
simulation program) from data processing (for statistical reports and
displays).  This can be very helpful in particular in large simulation 
programs or libraries, where different objects may need to process 
the same data in different ways.  
A statistical probe maintains a list of registered 
\externalclass{umontreal.iro.lecuyer.stat}{ObservationListener}
objects, and broadcasts information to all its registered
observers whenever appropriate.  %For the
%\externalclass{umontreal.iro.lecuyer.stat}{StatProbe} class,
%this happens whenever a new observation is given to the probe.
Any object that implements the interface 
\externalclass{umontreal.iro.lecuyer.stat}{ObservationListener}
can register as an observer. For an example, see the program 
\texttt{QueueObs} in the directory \texttt{examples}.

Subpackages of package \texttt{stat} provide matrices of \texttt{Tally}'s
(see the overview on page \pageref{over-mat-tal} of this guide),
and lists of \texttt{Tally}'s (see page \pageref{over-li-tal} of this guide).
% and lists of \texttt{Tally}'s with control variables
% (see page \pageref{over-licv-tal} of this guide).



%%%%%%%%%   Enlever ce qui suit: trop technique pour l'overview.
\begin{comment}

When writing complex simulation programs, in data generation portions,
one uses the statistical
probes as usual but with observation notification turned on.  To turn
on observation notification, one simply calls
\externalmethod{umontreal.iro.lecuyer.stat}{StatProbe}{notifyObs}{}
on the probes.  In this mode, a probe becomes a \emph{distribution
  agency} which broadcasts received observations to
all \emph{observers} that registered to it.  It will also fulfill its
collector functionality unless it is disabled explicitly by calling
the
\externalmethod{umontreal.iro.lecuyer.stat}{StatProbe}{stopCollectStat}{}
method.
When the statistical collection is disabled, the probes are only
distribution agencies.

Data processing parts of the program are implemented through
observers. An observer is a class
implementing the \externalclass{java.util}{Observer} interface.  It
can be registered to any class extending
\externalclass{java.util}{Observable}, including
\externalclass{umontreal.iro.lecuyer.stat}{StatProbe}.
The observed value is passed
to the observers through the second argument of the
\externalmethod{java.util}{Observer}{update}{} method which is an
\texttt{Object}.  The argument must then be
type-casted to a \texttt{Double} wrapper object
before the observation value can be extracted.

\end{comment}
