\defclass {NiedXingSequenceBase2}

This class implements digital sequences based on the
 Niederreiter-Xing sequence in base 2.
\latex{For details on these point sets, see \cite{rNIE98a,rPIR01c}.}
\hpierre{The current numbers used to fill up the generating matrices
   were taken from the web page of Gottlieb Pirsic some time ago
   and contain errors. We should check for an update on his web page.}
\hrichard{Les nouveaux nombres ont \'et\'e recopi\'es le 15 juillet 2004.}
\hrichard{Nouvelle r\'ef\'erence int\'eressante? \cite{rCLA99a}}

%%%%%%%%%%%%%%%%%%%%%%%

\bigskip\hrule
\begin{code}
\begin{hide}
/*
 * Class:        NiedXingSequenceBase2
 * Description:  Niederreiter-Xing sequences in base 2
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.hups; \begin{hide} 

import java.io.Serializable;
import java.io.ObjectInputStream;
import java.io.InputStream;
import java.io.FileNotFoundException;
import java.io.IOException;
import umontreal.iro.lecuyer.util.PrintfFormat;
\end{hide}

public class NiedXingSequenceBase2 extends DigitalSequenceBase2 \begin{hide} { 

   private static final int MAXDIM  = 32;  // Maximum dimension.
   private static final int NUMCOLS = 30;  // Maximum number of columns.
   private static final boolean isTrans = true;
\end{hide} 
\end{code}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection* {Constructors}
\begin{code} 

   public NiedXingSequenceBase2 (int k, int w, int dim) \begin{hide} {
      init (k, w, w, dim);
   } 
\end{hide}
\end{code} 
\begin{tabb}
    Constructs a new Niederreiter-Xing digital sequence in base 2 
    with $n = 2^k$ points and $w$ output digits, in \texttt{dim} dimensions.
    The generator matrices $\mathbf{C}_j$ are $w\times k$ and
    the numbers making the bit matrices are taken from
   \htmladdnormallink{Pirsic's site}{http://www.ricam.oeaw.ac.at/people/page/pirsic/niedxing/index.html}.
  The bit matrices from Pirsic's site are transposed to be consistent with SSJ,
  and at most 30 bits of the matrices are used. 
%    Unless, one plans to apply a randomization on more than $k$ digits
%     (e.g., a random digital shift for $w > k$ digits, or a linear
%     scramble yielding $r > k$ digits), one should 
%     take $w = r = k$ for better computational efficiency.
    Restrictions: $0\le k\le 30$, $k\le w$, and \texttt{dim} $\le 32$.
\end{tabb}
\begin{htmlonly}
   \param{k}{there will be 2^k points}
   \param{w}{number of output digits}
   \param{dim}{dimension of the point set}
\end{htmlonly}
\begin{code}\begin{hide}

   public String toString() {
      StringBuffer sb = new StringBuffer ("Niederreiter-Xing sequence:" +
                                           PrintfFormat.NEWLINE);
      sb.append (super.toString());
      return sb.toString();
   }

   private void init (int k, int r, int w, int dim) {
      if ((dim < 1) || (dim > MAXDIM))
         throw new IllegalArgumentException 
            ("Dimension for NiedXingSequenceBase2 must be > 1 and <= " + MAXDIM);
      if (r < k || w < r || w > MAXBITS || k >= MAXBITS) 
         throw new IllegalArgumentException
        ("One must have k < 31 and k <= r <= w <= 31 for NiedXingSequenceBase2");
      numCols   = k;
      numRows   = r;
      outDigits = w;
      numPoints = (1 << k);
      this.dim  = dim;
      // 1L otherwise gives wrong results for outDigits >= 31
      normFactor = 1.0 / ((double) (1L << (outDigits)));
      genMat = new int[dim * numCols];
      initGenMat();
   } 


   public void extendSequence (int k) {
      init (k, numRows, outDigits, dim);
   }


   // Initializes the generator matrices for a sequence.
   private void initGenMat ()  {
      // Compute where we should start reading matrices.
      /* Pirsic gives matrices starting at dimension 4; there are j matrices
         of dimension j. Thus if we want to use matrices in dimension
         dim, we must jump over all matrices of dimension < dim. If they
         started at dimension 1, there would be dim*(dim-1)/2 elements to
         disregard. But the first 3 dimensions are not there, thus subtract
         3*(3 - 1)/2 = 6 to get the starting element of dimension dim.
      I also multiply by 2 because the relevant columns are in the 30 least 
         significant bits of NiedXingMat, but if I understand correctly,
         SSJ assumes that they are in bits [31, ..., 1].
      At last, I shift right if w < 31 bits. */

      int start;
      if (dim <= 4) 
         start = 0;
      else
         start = ((dim*(dim-1)/2)-6)*NUMCOLS;

      long x;
      if (isTrans) {
         for (int j = 0; j < dim; j++)
            for (int c = 0; c < numCols; c++) {
               x = NiedXingMatTrans[start + j*NUMCOLS + c];
               x <<= 1;
               genMat[j*numCols + c] = (int) (x >> (MAXBITS - outDigits));
             }
      } else {
         for (int j = 0; j < dim; j++)
            for (int c = 0; c < numCols; c++) {
               x = NiedXingMat[start + j*NUMCOLS + c];
               x <<= 1;
               genMat[j*numCols + c] = (int) (x >> (MAXBITS - outDigits));
            }
      }
   }


   // ****************************************************************** 
   // Generator matrices of Niederreite-Xing sequences.
   // This array stores explicitly NUMCOLS columns in MAXDIM dimensions.
   // The implemented generator matrices are provided by Gottlieb Pirsic
   // and were downloaded on 15 July 2004. (RS)
   // These are the numbers given by Pirsic: I kept the first 30 of each
   // row vector in each dimension and shifted them to get the 30 most 
   // significant bits. These numbers considered as 30 X 30 bit matrices 
   // are given in matrix NiedXingMat below. The same numbers but transposed
   // as 30 X 30 bit matrices are given in matrix NiedXingMatTrans below.
   // According to Yves Edel, the correct matrices for Niederreiter-Xing
   // are NiedXingMatTrans.
   //
   // The matrices were given up to MAXDIM = 32, but the javac compiler
   // cannot compile code  that is too big. So I serialized them in
   // file NiedXingSequenceBase2.ser. (RS)

   private static int[] NiedXingMat;
   private static int[] NiedXingMatTrans;
   private static final int MAXN = 15660;

   static {
//      NiedXingMat = new int[MAXN];
      NiedXingMatTrans = new int[MAXN];

      try {
/*
         InputStream is = 
            NiedXingSequenceBase2.class.getClassLoader().getResourceAsStream (
         "umontreal/iro/lecuyer/hups/dataSer/Nieder/NiedXingSequenceBase2.ser");
         if (is == null)
            throw new FileNotFoundException (
               "Cannot find NiedXingSequenceBase2.ser");
         ObjectInputStream ois = new ObjectInputStream(is);
         NiedXingMat = (int[]) ois.readObject();
         ois.close();
*/
         InputStream is =
            NiedXingSequenceBase2.class.getClassLoader().getResourceAsStream(
      "umontreal/iro/lecuyer/hups/dataSer/Nieder/NiedXingSequenceBase2Trans.ser");
         if (is == null)
            throw new FileNotFoundException (
               "Cannot find NiedXingSequenceBase2Trans.ser");
         ObjectInputStream ois = new ObjectInputStream(is);
         NiedXingMatTrans = (int[]) ois.readObject();
         ois.close();

      } catch(FileNotFoundException e) {
         e.printStackTrace();
         System.exit(1);

      } catch(IOException e) {
         e.printStackTrace();
         System.exit(1);

      } catch(ClassNotFoundException e) {
         e.printStackTrace();
         System.exit(1);
      }
   }
}
\end{hide}
\end{code}

