\defclass{RandShiftedPointSet}

\adam{Cette classe reprogramme un \texttt{addRandomShift} explicitement sur
l'ensemble $P$, alors
que le \texttt{ContainerPointSet} applique le \texttt{addRandomShift} sur le
contenu $P$,  ce qui est beaucoup plus propre et plus g\'en\'eral.
Faut-il \'eliminer cette classe? Est-elle vraiment n\'ecessaire?}
%
This container class embodies a point set to which a random shift
modulo 1 is applied (i.e., a single uniform random point is added
to all points, modulo 1, to randomize the inner point set).
When calling \method{addRandomShift}{}, a new random shift will be generated.
This shift is represented by a vector of $d$ uniforms over $(0,1)$,
where $d$ is the current dimension of the shift.

% The shift is executed \emph{after} all other randomizations in the
% randomization list, if any.


\bigskip\hrule\bigskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{code}
\begin{hide}
/*
 * Class:        RandShiftedPointSet
 * Description:  Point set to which a random shift modulo 1 is applied
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       
 * @since

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.hups;\begin{hide}

import umontreal.iro.lecuyer.util.PrintfFormat;
import umontreal.iro.lecuyer.rng.*;
\end{hide}

public class RandShiftedPointSet extends ContainerPointSet \begin{hide} {

   protected double[] shift;           // The random shift.
   protected int dimShift = 0;         // Current dimension of the shift.
   protected int capacityShift = 0;    // Number of array elements;
                                       // always >= dimShift.
   protected RandomStream shiftStream; // Used to generate random shifts.
\end{hide}
\end{code}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Constructor}

\begin{code}

   public RandShiftedPointSet (PointSet P, int dimShift, RandomStream stream)\begin{hide} {
      init (P);
      if (dimShift <= 0) {
         throw new IllegalArgumentException (
            "Cannot construct RandShiftedPointSet with dimShift <= 0");
      }
      shiftStream = stream;
      shift = new double [dimShift];
      capacityShift = this.dimShift = dimShift;
   }\end{hide}
\end{code}
 \begin{tabb}
  Constructs a structure to contain a randomly shifted version of \texttt{P}.
  The random shifts will be generated in up to \texttt{dimShift} dimensions,
  using stream \texttt{stream}.
 \end{tabb}
\begin{htmlonly}
   \param{P}{point set being randomized}
   \param{dimShift}{dimension of the initial shift}
   \param{stream}{stream used for generating random shifts}
\end{htmlonly}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection*{Methods}
\begin{code}

   public int getShiftDimension()\begin{hide} {
      return dimShift;
   }\end{hide}
\end{code}
 \begin{tabb}
  Returns the number of dimensions of the current random shift.
 \end{tabb}
\begin{code}

   public void addRandomShift (int d1, int d2, RandomStream stream) \begin{hide} {
      if (null == stream)
         throw new IllegalArgumentException (
              PrintfFormat.NEWLINE +
                  "   Calling addRandomShift with null stream");
      if (stream != shiftStream)
         shiftStream = stream;
      addRandomShift (d1, d2);
   }\end{hide}
\end{code}
 \begin{tabb}
  Changes the stream used for the random shifts to \texttt{stream}, then
  refreshes the shift for coordinates \texttt{d1} to \texttt{d2-1}.
\richard{Il y a 4 m\'ethodes \texttt{addRandomShift}. Peut-\^etre faudrait-il
en \'eliminer 2, comme dans \texttt{PointSet}.}
 \end{tabb}
\begin{code}

   public void addRandomShift (RandomStream stream) \begin{hide} {
      if (stream != shiftStream)
         shiftStream = stream;
      addRandomShift (0, dimShift);
   }\end{hide}
\end{code}
 \begin{tabb}
  Changes the stream used for the random shifts to \texttt{stream}, then
  refreshes all coordinates of the random shift, up to its current dimension.
 \end{tabb}
\begin{hide}
\begin{code}

   @Deprecated
   public void addRandomShift (int d1, int d2)\begin{hide} {
      if (d1 < 0 || d1 > d2)
         throw new IllegalArgumentException ("illegal parameter d1 or d2");
      if (d2 > capacityShift) {
         int d3 = Math.max (4, capacityShift);
         while (d2 > d3)
            d3 *= 2;
         double[] temp = new double[d3];
         capacityShift = d3;
         for (int i = 0; i < d1; i++)
            temp[i] = shift[i];
         shift = temp;
      }
      dimShift = d2;
      for (int i = d1; i < d2; i++)
         shift[i] = shiftStream.nextDouble();

 // Just for testing, to see the single uniform random point
 //     for(int k = 0; k < d2; k++)
 //       System.out.println ("shift " + k + " = " + shift[k]);
 //     System.out.println();

   }\end{hide}
\end{code}
 \begin{tabb}
  Refreshes the random shift (generates new uniform values for the
  random shift coordinates) for coordinates \texttt{d1} to \texttt{d2-1}.
 \end{tabb}
\begin{code}

   @Deprecated
   public void addRandomShift() \begin{hide} {
      addRandomShift (0, dimShift);
   }\end{hide}
\end{code}
 \begin{tabb}
  Refreshes all coordinates of the random shift, up to its current dimension.
 \end{tabb}
\end{hide}
\begin{code}\begin{hide}

   public String toString() {
      return "RandShiftedPointSet of: {" + PrintfFormat.NEWLINE
              + P.toString() + PrintfFormat.NEWLINE + "}";
   }

   public PointSetIterator iterator() {
      return new RandShiftedPointSetIterator();
   }

   // ***************************************************************

   private class RandShiftedPointSetIterator
                 extends ContainerPointSetIterator {

      public double getCoordinate (int i, int j) {
         int d1 = innerIterator.getCurCoordIndex();
         if (dimShift <= d1)
            // Must extend randomization.
            addRandomShift (dimShift, 1 + d1);
         double u = P.getCoordinate(i, j) + shift[j];
         if (u >= 1.0)
            u -= 1.0;
         if (u > 0.0)
            return u;
         return EpsilonHalf;  // avoid u = 0
      }

      public double nextCoordinate() {
         int d1 = innerIterator.getCurCoordIndex();
         if (dimShift <= d1)
            addRandomShift (dimShift,1 + d1 );
         double u = shift [d1];
         u += innerIterator.nextCoordinate();
         if (u >= 1.0)
            u -= 1.0;
         if (u > 0.0)
            return u;
         return EpsilonHalf;  // avoid u = 0
      }

   }
}\end{hide}
\end{code}
