\defclass{IIDMultivariateGen}

Extends \class{RandomMultivariateGen} for a vector of independent identically distributed
(i.i.d.) random variables.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bigskip\hrule

\begin{code}
\begin{hide}
/*
 * Class:        IIDMultivariateGen
 * Description:  vector of independent identically distributed random variables
 * Environment:  Java
 * Software:     SSJ 
 * Copyright (C) 2001  Pierre L'Ecuyer and Université de Montréal
 * Organization: DIRO, Université de Montréal
 * @author       David Munger
 * @since        January 2011

 * SSJ is free software: you can redistribute it and/or modify it under
 * the terms of the GNU General Public License (GPL) as published by the
 * Free Software Foundation, either version 3 of the License, or
 * any later version.

 * SSJ is distributed in the hope that it will be useful,
 * but WITHOUT ANY WARRANTY; without even the implied warranty of
 * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
 * GNU General Public License for more details.

 * A copy of the GNU General Public License is available at
   <a href="http://www.gnu.org/licenses">GPL licence site</a>.
 */
\end{hide}
package umontreal.iro.lecuyer.randvarmulti;
\begin{hide}
import umontreal.iro.lecuyer.randvar.RandomVariateGen;
\end{hide}

public class IIDMultivariateGen extends RandomMultivariateGen\begin{hide} {
\end{hide}\end{code}

\subsubsection*{Constructor}

\begin{code}

   public IIDMultivariateGen (RandomVariateGen gen1, int d) \begin{hide} {
      setGen1 (gen1);
      this.stream = gen1.getStream();
      dimension = d;
   } \end{hide}
\end{code}
\begin{tabb}
Constructs a generator for a \texttt{d}-dimensional vector of i.i.d.\ variates
with a common one-dimensional generator \texttt{gen1}.
\end{tabb}
\begin{htmlonly}
   \param{gen1}{the one-dimensional generator}
   \param{d}{dimension of the vector (number of i.i.d.\ variates).}
\end{htmlonly}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\subsubsection*{Methods}

\begin{code}

   public void setDimension (int d) \begin{hide} {
      dimension = d;
   } \end{hide}
\end{code}
\begin{tabb}
Changes the dimension of the vector to \texttt{d}.
\end{tabb}
\begin{code}

   public void nextPoint (double[] p) \begin{hide} {
      if (p.length != dimension)
         throw new IllegalArgumentException(String.format(
            "p's dimension (%d) does not mach dimension (%d)", p.length, dimension));

      for (int i = 0; i < dimension; i++)
         p[i] = gen1.nextDouble();
   } \end{hide}
\end{code}
\begin{tabb}
Generates a vector of i.i.d.\ variates.
\end{tabb}
\begin{code}

   public void setGen1 (RandomVariateGen gen1)\begin{hide} {
      if (gen1 == null)
         throw new NullPointerException ("gen1 is null");
      this.gen1 = gen1;
   } \end{hide}
\end{code}
\begin{tabb}
Sets the common one-dimensional generator to \texttt{gen1}.
\end{tabb}
\begin{code}

   public RandomVariateGen getGen1()\begin{hide} {
     return gen1;
   } \end{hide}
\end{code}
\begin{tabb}
Returns the common one-dimensional generator used in this class.
\end{tabb}
\begin{code}

   public String toString() \begin{hide} {
      return dimension + "-dimensional vector of i.i.d. " +
            gen1.getDistribution().toString();
   }
   \end{hide}
\end{code}
\begin{tabb}
Returns a string representation of the generator.
\end{tabb}

\begin{code}\begin{hide}
}
\end{hide}\end{code}
