\documentclass[12pt]{article}
\usepackage{html}
\usepackage{url}
\usepackage{ssj}

\mytwoheads
\dateheadtrue
\detailedtrue

\begin{document}

%begin{latexonly}
\begin{titlepage}

\null\vfill
\begin {center}
%{\Large\bf SSJ User's Guide } \\[20pt] 
{\Large SSJ: Stochastic Simulation in Java}\\[15pt]
{\Large Overview} \\[20pt]
 Version: \today \\
\vfill
\end {center}
%end{latexonly}

SSJ is a Java library for stochastic simulation, developed in the 
% simulation laboratory of the 
D\'epartement d'Informa\-tique et de Recherche
Op\'erationnelle (DIRO), at the Universit\'e de Montr\'eal.
It provides facilities for generating uniform and nonuniform random 
variates, computing different measures related to probability 
distributions, performing goodness-of-fit tests, applying 
quasi-Monte Carlo methods, collecting statistics (elementary),
and programming discrete-event simulations with both events and processes.
Additional Java packages are also developed on top of SSJ for simulation
applications in finance, call centers management, communication networks,
 etc.
\bigskip


\begin{center}
\fbox{\parbox{0.94\textwidth}{ 
    \ttfamily
    This file is part of SSJ.

    Copyright (C) 2008  Pierre L'Ecuyer and Universit\'e de Montr\'eal
\medskip

    SSJ is free software: you can redistribute it and/or modify
    it under the terms of the GNU General Public License as published by
    the Free Software Foundation, either version 3 of the License, or
    any later version.
\medskip

    SSJ is distributed in the hope that it will be useful,
    but WITHOUT ANY WARRANTY; without even the implied warranty of
    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
    GNU General Public License for more details.
\medskip

    You should have received a copy of the GNU General Public License
    along with SSJ.  If not, see \url{http://www.gnu.org/licenses}.
}}
\end{center}
%begin{latexonly}
\vfill
\end{titlepage}
%end{latexonly}

%%%%%%%%%%%%%%
\subsection*{Introduction and overview}

Simulation models can be implemented in many ways \cite{sLAW00a}.
One can use general-purpose programming languages such as 
FORTRAN, C, C++, Java, or specialized simulation languages such as
\emph{GPSS}, \emph{SIMAN}, and \emph{SIMSCRIPT}.
The general-purpose languages may be
more familiar to the programmer, but usually
do not have the necessary built-in tools to perform simulation.
Implementing a model can become complex and tedious.  
Specialized simulation languages must be learned before models 
can be implemented, and they are not as widely available and supported
as the most popular general-purpose languages.

Over the past few decades, commercial simulation tools with point-and-click 
graphical user interfaces such as \emph{Arena}, \emph{Automod},
 \emph{Witness}, 
and many others, have become by far the most widely used tools to develop
simulation models.  Among their main advantages, these tools do not require
knowledge of a programming language, provide graphical animation,
have automatic facilities to collect statistics and perform experiments,
and can sometimes perform optimization to a certain extent.
On the other hand, these specialized simulation tools,
especially the point-and-click tools, are often too restrictive, 
because they are usually targeted at a limited class of models.
With these tools, simulating a system whose logic is complicated or 
unconventional may become quite difficult.
All the graphical and automatic devices also tend to slow down the simulation
significantly.  Fast execution times are important for example in a context
of optimization, where thousands of variants of a base system may have to be
simulated, or for on-line applications where a fast response time is required.


SSJ is an organized set of packages whose purpose is to facilitate 
simulation programming in the Java language.
A first description was given in \cite{sLEC02a}.
Some of the tools can also be used for modeling (e.g., selecting
and fitting distributions).
\begin{comment}
A simulation library, such
as SSJ, extends the Java programming language to equip it with the
necessary tools for implementing complex models.  One can use a
familiar programming language, such as Java, with high-level tools for
simulation.  A simulation project can be divided in several tasks,
from modeling to implementation \cite{sLAW00a}.  SSJ provides tools
for model implementation and validation.
\end{comment}
As these lines are being written, SSJ is still growing.
Several new packages, classes, and methods will certainly be added in 
forthcoming years and others will be refined.

The facilities offered are grouped into different packages, 
each one having its own user's guide, in the form of a PDF file.
There is also a set of commented examples of simulation programs
in a separate directory with its own guide.
Programs are given for some of the examples used in the books of 
Law and Kelton \cite{sLAW00a} and Glasserman \cite{fGLA04a}, for instance.
The best way to learn about SSJ, at the beginning, is probably
to study these examples and refer to the user guides of the different
packages when needed.
The PDF files are the official documentation.
There is also a simplified on-line documentation in HTML 
format, produced via \texttt{javadoc}.


\input{packdep.tex}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{comment}

% Three packages are used for uniform and non-uniform random number
% generation, one for the uniform random number generators, one for the
% probability distributions and one for the non-uniform random number
% generation.  It is also possible to replace the streams of uniform
% random numbers with highly uniform point sets for Quasi-Monte Carlo
% \cite{vLEC02a} simulation.
% As opposed to most random number generation libraries, it
% is also possible to compute densities, distribution functions and
% inverse distribution functions for all supported distributions.
% The \externalclass{umontreal.iro.lecuyer}{simevents}
% and \externalclass{umontreal.iro.lecuyer}{simprocs} packages make the
% heart of the SSJ library.
% They provide an efficient framework for stochastic simulation,
% supporting the event view, proces view, continuous simulation, and arbitrary 
% mixtures of these.
% The other packages allow one to perform miscellaneous tasks, such as
% statistical collection and goodness of fit tests.

\subsection*{random number generation}

Random numbers feed simulation models and allow one to compute
statistics.  To generate random numbers from
any probability distribution, uniform random numbers are required.
Such numbers are uniformly distributed in the $[0,1)$ interval, i.e.,
the probability of getting a given number $x$ in that interval is the same
for all values of $x\in[0,1)$.  Any generated number $x$ is also
independent from any previous or future generated numbers.  Although
the generated uniforms are not truly independent since one uniform is
obtained from the previous uniforms by a mathematical formula, one can
consider them independent for simulation purposes.  Selection of a
random number generator is based on several criteria such as
uniformity, performance, and portability \cite{rLEC01d}.
The package
\externalclass{umontreal.iro.lecuyer}{rng} contains the needed tools
to generate such numbers.  It defines an interface called
\externalclass{umontreal.iro.lecuyer.rng}{RandomStream} implemented by
any random number generator supported by SSJ.  This interface allows
one to easily interchange random number generators since they are
accessed through the same set of methods specified by the interface.
Only the random number generator setup depends on the type of
generator that was chosen.

If one wants to replace uniform random numbers with low-discrepancy
point sets for variance reduction, the package
\externalclass{umontreal.iro.lecuyer}{hups} contains all the necessary
facilities.  Such highly uniform point sets all inherit from the
\externalclass{umontreal.iro.lecuyer.hups}{PointSet} which provides a
\externalclass{umontreal.iro.lecuyer.hups}{PointSetIterator} extending
\externalclass{umontreal.iro.lecuyer.rng}{RandomStream}.  The
replacement can be easily done without modifying the model
implementation, except the setup-time code.

To generate non-uniform random numbers, one must select a probability
distribution based on the empirical data \cite{sLAW00a}.
SSJ does not provide
probability distribution estimation tools, but goodness of fit tests are
included to help in model validation.  The package
\externalclass{umontreal.iro.lecuyer}{probdist} contains several
standard, commonly-used, probability distributions.  It supports
discrete and continuous distributions through two different abstract
base classes:
\externalclass{umontreal.iro.lecuyer.probdist}{ContinuousDistribution}
and
\externalclass{umontreal.iro.lecuyer.probdist}{DiscreteDistribution},
respectively.  Again, since the distributions inherit from a common
class, their access can be independent from the selected distribution,
except for the setup case.  One can compute the density/mass, distribution,
complementary, and inverse distribution functions.  These facilities
are also accessible through static methods implemented in each
distribution class if one does not want to create objects or needs
distributions whose parameters vary in time.
However, setup-time operations must be performed for each operation,
which can be inefficient for certain distributions.

To generate non-uniform random numbers, the packages
\externalclass{umontreal.iro.lecuyer}{rng} (or
\externalclass{umontreal.iro.lecuyer}{hups}) and
\externalclass{umontreal.iro.lecuyer}{probdist} must be used
together.  The simplest generation method is to generate a uniform
random number using a generator implementing
\externalclass{umontreal.iro.lecuyer.rng}{RandomStream} (or get a
coordinate using a point set iterator) and to apply
inversion by using the selected
\externalclass{umontreal.iro.lecuyer}{probdist} distribution's
\externalmethod{umontreal.iro.lecuyer.probdist}{ContinuousDistribution}{inverseF}{} method.
However, inversion is not the only generation method and sometimes not
the most efficient.  For some distributions, closed-form inverse functions
or fast inversion algorithms exist.
For others, inversion is performed using
binary or even linear search.  In such cases, the performance
and precision depends on the complexity of the distribution function
which is calculated several times for one inverse.
The package
\externalclass{umontreal.iro.lecuyer}{randvar} acts as glue between
uniform random number generators and probability distributions.  Continuous
or discrete random number generators also inherits from common base
classes, namely
\externalclass{umontreal.iro.lecuyer.randvar}{RandomVariateGen} and
\externalclass{umontreal.iro.lecuyer.randvar}{RandomVariateGenInt}.
All generators use a random stream and a probability
distribution for their construction.  As opposed to
\externalclass{umontreal.iro.lecuyer}{probdist}, one can directly
instantiate
\externalclass{umontreal.iro.lecuyer.randvar}{RandomVariateGen} or
\externalclass{umontreal.iro.lecuyer.randvar}{RandomVariateGenInt}.
However, in such cases, only inversion generation method will be
available.  To use an alternate generation method, one must
instantiate a specialized generator class and switch to the given
generation algorithm using an object method.
Each specialized class also provides static method which perform
the same action.  Although they allow one to avoid object creation,
their signatures are specific to the used distribution and they have
to perform setup-time operations on each variate generation, which
can become inefficient.
The \externalclass{umontreal.iro.lecuyer}{randvar} package also
provides the class
\externalclass{umontreal.iro.lecuyer.randvar}{RandomVariateTrans} to
apply transformations to a generator.  This allows, for example, to
generate variates from a truncated distribution.

\subsection*{Performing simulation}

SSJ supports discrete-event, process-driven, continuous or mixed
simulation.  The discrete-event and continuous simulation are managed
by the package \externalclass{umontreal.iro.lecuyer}{simevents}.  This
package manages the simulation clock and the event list, two essential
components for all discrete-event simulations.  The simulation clock
tracks the simulation time whereas the event list stores the
scheduled events to execute them in the right order.
Events are user-defined subclasses of
\externalclass{umontreal.iro.lecuyer.simevents}{Event}.  When an event
occurs, any type of actions can then be taken.  The package
provides a class called
\externalclass{umontreal.iro.lecuyer.simevents}{List} which implements
a linked list supporting statistical collection.  Continuous
simulation can be performed using the class
\externalclass{umontreal.iro.lecuyer.simevents}{Continuous}.  It uses
the event framework to resolve differential equations numerically at
fixed steps in the simulation time.

Process-driven simulation requires a separate package called
\externalclass{umontreal.iro.lecuyer}{simprocs}.  This package
provides the base class
\externalclass{umontreal.iro.lecuyer.simprocs}{SimProcess} which must
be extended by any process.  A simulation process is an autonomous
object which executes tasks and interacts with other simulation
processes.  To allow such interactions, the package provides some
synchronization tools such as a resource, a bin, and a condition.
The process-driven simulation framework relies on the event-driven
framework.  Processes start, resume and wake up on scheduled events.
One can then easily mix event-driven and process-driven simulation.

\subsection*{Other tools}

The package \externalclass{umontreal.iro.lecuyer}{stat} provides basic
tools for statistical collection.  Statistics are collected using
statistical probes, i.e, objects implementing the abstract class
\externalclass{umontreal.iro.lecuyer.stat}{StatProbe}.  Two types of
probes are supported.  The
\externalclass{umontreal.iro.lecuyer.stat}{Tally} allows to collect
observations of the form $X_1,\dots,X_n$ whereas
\externalclass{umontreal.iro.lecuyer.simevents}{Accumulate} collects
statistics for a continuous variable evolving in simulation time.
During the simulation, one can add observations to such probes.  After
the simulation, measures can be obtained, such as sample average,
sample standard deviation or confidence interval.  A statistical
report can be obtained for all probes.  The package also provides
a way to detach statistical collection from the model implementation
by using bound properties.

To test a proposed model against empirical data, goodness of fit tests
are provided in the package
\externalclass{umontreal.iro.lecuyer}{gof}.  
Such tests, e.g.\ Kolmogorov-Smirnov
or Anderson-Darling, compute a statistic using the
empirical observations and the proposed distribution.  The empirical
observations are given as an array whereas the
distribution is given as a
\externalclass{umontreal.iro.lecuyer}{probdist} object.  From the computed
statistic, it is possible to compute the $p$-value which is useful to
evaluate the significance of the test.

\subsection*{Related documentation}

The \texttt{example.pdf} file, in the \texttt{doc/pdf} subdirectory of
the SSJ distribution, explains simulation examples implemented using
SSJ.  This may be the best starting point to learn SSJ.
\begin{htmlonly}
% This should be removed if we find a way to display blbliographical
% references in HTML
% LaTeX2HTML can do it when it is run for a LaTeX document, but it does not
% work for a single bibliography shared by several documents.
% Texjava processes each Java class as a separate, standalone, LaTeX document.
One can find additional information and references in the PDF version
of this documentation, available in the {\tt doc/pdf} subdirectory
of the SSJ distribution.
\end{htmlonly}
\begin{latexonly}
% In Javadoc, this is visible to the user as hyperlinks.  In LaTeX,
% this is separated filed.
Every package introduced here contains its own reference documentation
as a PDF file, in the \texttt{doc/pdf} subdirectory.
This documentation describes in more details how to
use the package and provides a description of each class and method.
\end{latexonly}


\end{comment}
%%%%%%%%%%%%%%%%%%%%%%

%For overviews, tutorials, examples, guides, and tool documentation, 
%    please see: 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Acknowledgments}

SSJ was designed and implemented under the supervision of 
Pierre L'Ecuyer, with the contribution of the following persons
\begin{verse}
Mathieu Bague,
Sylvain Bonnet,
 \'Eric Buist, 
Maxime Dion,
 Yves Edel,
 Regina H.{} S.{} Hong, 
 Alexander Keller,
 Pierre L'Ecuyer, 
 \'Etienne Marcotte,
 Lakhdar Meliani, 
 Fran\c{c}ois Panneton,
Jean-Sebastien Parent-Chartier,
 Richard Simard,
 Cl\'ement Teule,
  Pierre-Alexandre Tremblay, 
 Jean Vaucher.
\end{verse}

Its development has been supported by NSERC-Canada grant No.\ ODGP0110050,
NATEQ-Qu\'ebec grant No.\ 02ER3218, a Killam fellowship,
and a Canada Research Chair to the author.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%begin{latexonly}
\bibliographystyle{plain}
\bibliography{simul,ift,random,fin}
%end{latexonly}

\end{document}
